---
title: "Summary statistics Production data"
author: "Luna Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
   # theme: cosmo
   # highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F,
                      root.dir = "e:/500390_OAD/output/sumstat_production/" )
```

Preliminary analysis of **ProductionExtract.csv**  

The objective is to get cleaned useful herds and cows for subsequent analysis.
Also to generate milk type (OAD, TAD, Transition) and find the transition year.

```{r, include=F}
# Luna Zhang Feb 28 2022

# library(data.table)
suppressWarnings({
  library(tidyverse)
  library(ggpubr)
})

# source("utils.R")

fileDir <- "e:/500390_OAD/output/" # "z:/Luna/500390_OAD/explore/"
rawfileDir <- "z:/Luna/500390_OAD/OAD Dairy Data/"
outputDir <- "e:/500390_OAD/output/sumstat_production/" # copy to "z:/Luna/500390_OAD/output/" later

if(!exists(outputDir)) {
  dir.create(outputDir)
}
# stop()

df_new <- readRDS(paste0(fileDir, 
                         "production_after2006_noTransferHerd_noGapYearHerd_",
                         "noGapYearAnim_noUnclassifiedSeasonMilkingType_",
                         "noBackToTADHerd.RData"))
```

```{r, eval = F}
nrow(df_new)
unique(df_new$HerdDurableKey) %>% length() #
unique(df_new$HerdDurableKey) %>% length() #
unique(df_new$AnimalDurableCode) %>% length()

names(df_new)
# [1] "AnimalDurableCode"             "DaysFromBirthToParturition"   
# [3] "DaysFromParturitionToHerdTest" "InducedParturitionFlag"       
# [5] "Volume_ml"                     "FatVolume_ml"                 
# [7] "ProteinVolume_ml"              "AgeParity"                    
# [9] "HerdDurableKey"                "DairyYear"                    
# [11] "Season"                        "EventDate"                    
# [13] "AgeParityGroup"                "RegimeType"                   
# [15] "event_date"                    "year"      
```

## Year and season count

2007 and 2022 season 0 have less data points.

```{r, fig.height=10}
year_season <- select(df_new, HerdDurableKey, AnimalDurableCode, year, Season) %>% 
  mutate(year_season = paste(year, Season, sep = "_")) 

cross_cow_year_season <- group_by(year_season, year_season) %>% 
  tally()
# print(cross_cow_year_season)
ggbarplot(cross_cow_year_season, "year_season", "n", 
          orientation = "horiz", fill = "steelblue",
          label = T, label.pos = "out")

# sumstat_year_season <- group_by(year_season, HerdDurableKey, year_season) %>% 
#   tally()
```

## Herd test distribution

Across herds and animals number of herd tests
```{r}
# within herd n herd tests
n_herd_tests <- group_by(df_new, HerdDurableKey, AnimalDurableCode) %>% 
  tally()

sumstat_n_tests <- group_by(n_herd_tests, HerdDurableKey) %>% 
  summarise(mean = mean(n), median = median(n), min = min(n), max = max(n)) %>% 
  pivot_longer(-HerdDurableKey, names_to = "statistics", values_to = "value")

group_by(df_new, AnimalDurableCode) %>%
  tally() %>%
  gghistogram(x = "n", binwidth = 1, add = "mean", fill = "steelblue",
              title = paste0("Numer of herd tests per cow, ", 
                             length(unique(df_new$AnimalDurableCode)), 
                             " total cows"))
```

Within herds number of herd tests

```{r}
gghistogram(sumstat_n_tests, x = "value", facet.by = "statistics", rug = T,
            add = "mean", binwidth = 1, fill = "steelblue",
            title = paste0("Number of herd tests per cow in each herd, ", 
                           nrow(sumstat_n_tests)/4, " total herds"))

```

## Age Parity distribution

Across herds and animal
```{r pressure}
# ageparity distr across herd vs herd mean distr
min_year <- min(df_new$year)
max_year <- max(df_new$year)

# 2007 SEASON 0 and 2006 season 1 are the same season
test <- filter(df_new, year==min_year & Season == 1)
season_2007 <- filter(df_new, year==min_year+1 & Season==0)
season_2007 <- rbind(test, season_2007) # contains multiple tests of one cow

# Find the index of 1st occurrence of each animal ID
a <- data.frame(table(season_2007$AnimalDurableCode))
a$var <- as.character(a$Var1)
# all(a$var==unique(season_2007$AnimalDurableCode)) # sanity check
a <- a[match(unique(season_2007$AnimalDurableCode), a$var),]
# all(a$var==unique(season_2007$AnimalDurableCode)) # sanity check
a$idx <- cumsum(a$Freq)-a$Freq + 1
idx <- a$idx
idx <- c(idx, nrow(season_2007)+1) # for the last set of animal IDs◘
a <- NULL

# length(idx)-1 # 490300 unique animal ID
idx <- tail(idx-1, -1) # last record of each animal ID
season_2007 <- season_2007[idx,]
# length(unique(season_2007$AnimalDurableCode))==nrow(season_2007) # sanity check

sumstat_ageparity_herd_2007 <- group_by(season_2007, HerdDurableKey) %>% 
  summarise(mean = mean(AgeParity), median = median(AgeParity),
            min = min(AgeParity), max = max(AgeParity)) %>% 
  pivot_longer(-HerdDurableKey, names_to = "statistics", values_to = "value")

summarise(season_2007, mean = mean(AgeParity), median = median(AgeParity),
            min = min(AgeParity), max = max(AgeParity))
```

Within herds

```{r}
gghistogram(sumstat_ageparity_herd_2007, 
            x = "value", facet.by = "statistics", rug = T,
            add = "mean", binwidth = 1,  fill = "steelblue",
            title = paste0("AgeParity of the last herd test in 2007, ", 
                           nrow(sumstat_ageparity_herd_2007)/4, " total herds"))

```

## milk_type (OAD, TAD, TRANSITION), herd percent OAD distribution

```{r}
milk_type_herd <- select(df_new, HerdDurableKey, event_date, year, 
                         herd_milk_type) %>% 
  distinct()
# which(duplicated(milk_type_herd[1:2,])) %>% length() # sanity check

milk_type_herd <- select(df_new, HerdDurableKey, herd_milk_type) %>% distinct()
# length(unique(milk_type_herd$HerdDurableKey))==nrow(milk_type_herd) # sanity check
sumstat_milk_type_herd <- group_by(milk_type_herd, herd_milk_type) %>% 
  tally() %>% mutate(prop = round(n/sum(n), digits = 2))

print(sumstat_milk_type_herd)

```

## Transition year distribution

```{r}
transition_year <- select(df_new, HerdDurableKey, transition_year, 
                          herd_milk_type) %>% 
  distinct() %>% 
  filter(herd_milk_type == "Transition")

gghistogram(transition_year, "transition_year", binwidth = 1, fill = "steelblue",
            title = paste0("Transition year distribution among ", 
                           nrow(transition_year), " herds"))
```
```{r, eval = F}
rmarkdown::render(output_dir = outputDir, 
                  input = rstudioapi::getSourceEditorContext()$path)
```