---
title: "Summary statistics milking types"
author: "Luna ZHnag"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
   # theme: cosmo
   # highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F,
                      root.dir = "e:/500390_OAD/output/sumstat_production/" )
```

The objective is to get cleaned useful herds and cows for subsequent analysis.
Also to generate milk type (OAD, TAD, Transition) and find the transition year.

```{r, include=F}
# Luna Zhang Feb 28 2022

# library(data.table)
suppressWarnings({
  library(tidyverse)
  library(ggpubr)
})

# source("utils.R")

fileDir <- "e:/500390_OAD/output/" # "z:/Luna/500390_OAD/explore/"
rawfileDir <- "z:/Luna/500390_OAD/OAD Dairy Data/"
outputDir <- "e:/500390_OAD/output/sumstat_production/" # copy to "z:/Luna/500390_OAD/output/" later

if(!exists(outputDir)) {
  dir.create(outputDir)
}
# stop()

df_new <- readRDS(paste0(fileDir, 
                         "production_after2006_noTransferHerd_noGapYearHerd_",
                         "noGapYearAnim_noUnclassifiedSeasonMilkingType_",
                         "noBackToTADHerd.RData"))
event_milk_type <- read.csv(paste0(fileDir, "event_milk_type.csv"))
season_milk_type <- read.csv(paste0(fileDir, "season_milk_type.csv"))
names(event_milk_type)[1] <- "HerdDurableKey"
names(season_milk_type)[1] <- "HerdDurableKey"
herd_milk_type <- read.csv(paste0(fileDir, "herd_milk_type.csv"))
all_milk_type <- read.csv(paste0(fileDir, "herd_season_testday_milk_type.csv"))

```


## Milking type frequency at test day

For each Herd-Season-EventDate, define the milking type as 95% of the animals.

* $>=$ 95% are OAD, define as OAD. Same for TAD.
* If (5%, 95%) are TAD then define it as unknown. 
* If (5%, 95%) are OAD define it as unknown.

```{r}

knitr::kable(data.frame(event_milk_type = c("OAD", "TAD", "unknown"),
                        n = c(51835, 264979, 24101),
                        prop = c(0.15, 0.777, 0.071)),
             digits = 2, caption = "Before cleaning")

knitr::kable(group_by(event_milk_type, event_milk_type) %>% 
               tally() %>% 
               mutate(prop = n/sum(n)), digits = 2, caption = "After cleaning")

```

## Milking type frequency at each season

For herds that are not all OAD or all TAD in all test days:

Use January 1st as the start of the 2nd season (June 1 to Dec 1 is the 1st season)

* If all TAD before Jan 1st and all OAD after Jan 1st --> TAD, switch_type --> switch
* If OAD happened before Jan 1st --> unclassified, switch_type --> unknown
* If TAD happened after OAD --> unclassified, switch_type --> unknown
* If unknown occurs before Jan 1st --> unclassified, switch --> unknown

added later:

* Mixed of TAD and unknown (after Jan 1st) --> TAD, switch --> unknown
* Mixed of OAD and unknown (after Jan 1st) --> OAD, switch --> unknown

```{r}

knitr::kable(data.frame(season_milk_type = c("OAD", "TAD", "unclassified"),
                        n = c(6436, 82939, 9678),
                        prop = c(0.06497532, 0.83731941, 0.09770527)),
             digits = 2, caption = "Before cleaning")

knitr::kable(group_by(season_milk_type, season_milk_type, switch_type) %>% 
               tally() %>% 
               mutate(n/sum(n)), digits = 2, caption = "After cleaning")

```

## Milking type by herds

* All OAD across seasons --> OAD, same for TAD
* TAD to OAD --> transition
* TAD after OAD --> illegal, delete this herd
* If there is an unclassified, consider it as a gap year --> illegal,
delete this herd

```{r}
knitr::kable(data.frame(herd_milk_type = c("OAD", "TAD", "Transition", "illegal"),
                        n = c(530, 9274, 339, 413)) %>% 
               mutate(prop = n/sum(n)),
             digits = 2, caption = "Before cleaning")

knitr::kable(group_by(herd_milk_type, herd_milk_type) %>% 
               tally() %>% 
               mutate(n/sum(n)), digits = 2, caption = "After cleaning")

```

## combined together

```{r}
group_by(all_milk_type, herd_milk_type, season_milk_type, switch_type,
         event_milk_type) %>% 
  tally() %>% 
  mutate(prop = n/sum(n)) %>% 
  knitr::kable(digits = 2)

```

```{r, eval = F}
rmarkdown::render(output_dir = outputDir, 
                  input = rstudioapi::getSourceEditorContext()$path)
```
