---
title: "Find TAD herd outliers"
author: "Luna ZHnag"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F,
                      root.dir = "e:/500390_OAD/output/sumstat_production/" )
```


```{r, include=F}
# Luna Zhang Feb 28 2022

# library(data.table)
suppressWarnings({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(ggpubr)
})

# source("utils.R")

fileDir <- "e:/500390_OAD/output/" # "z:/Luna/500390_OAD/explore/"
rawfileDir <- "z:/Luna/500390_OAD/OAD Dairy Data/"
outputDir <- "C:/Users/lzhang/Documents/GitHub/500390_OAD/output/" 

if(!exists(outputDir)) {
  dir.create(outputDir)
}
# stop()
```

```{r, include = F}
full_df <- readRDS(paste0(fileDir,
                          "production_survival_noYoungAnimal_gebv_beforeParity5.RData"))
```

```{r, eval = F}
names(full_df)
#  [1] "AnimalDurableCode"             "DaysFromBirthToParturition"    "DaysFromParturitionToHerdTest"
#  [4] "InducedParturitionFlag"        "Volume_ml"                     "FatVolume_ml"                 
#  [7] "ProteinVolume_ml"              "AgeParity"                     "HerdDurableKey"               
# [10] "DairyYear"                     "Season"                        "EventDate"                    
# [13] "AgeParityGroup"                "RegimeType"                    "event_date"                   
# [16] "herd_milk_type"                "transition_year"               "s12"                          
# [19] "s23"                           "s34"                           "s45"                          
# [22] "EBV_BodyConditionScore"        "EBV_Fertility_CR42"            "EBV_Fertility_CSD0"           
# [25] "EBV_Fertility_CSD123"          "EBV_Fertility_PM123"           "EBV_FunctionalSurvival"       
# [28] "EBV_IndirectSurvival"          "EBV_SomaticCellScore"          "EBV_Survival"                 
# [31] "EBV_TOP_AdaptabilityToMilking" "EBV_TOP_Capacity"              "EBV_TOP_DairyConformation"    
# [34] "EBV_TOP_ForeUdder"             "EBV_TOP_FrontTeat"             "EBV_TOP_Legs"                 
# [37] "EBV_TOP_MilkingSpeed"          "EBV_TOP_OverallOpinion"        "EBV_TOP_RearTeat"             
# [40] "EBV_TOP_RearUdder"             "EBV_TOP_RumpAngle"             "EBV_TOP_RumpWidth"            
# [43] "EBV_TOP_ShedTemperament"       "EBV_TOP_Stature"               "EBV_TOP_TeatLength"           
# [46] "EBV_TOP_UdderOverall"          "EBV_TOP_UdderSupport"          "EBV_TotalFat"                 
# [49] "EBV_TotalProtein"              "EBV_TotalVolume"               "transition_parity"            
# [52] "EBV_Liveweight"                "GBV_TotalFat"                  "GBV_TotalProtein"             
# [55] "GBV_TotalVolume"               "GBV_Liveweight"                "GBV_SomaticCellScore"         
# [58] "GBV_Fertility_CR42"            "GBV_BodyConditionScore"        "GBV_FunctionalSurvival"       
# [61] "GBV_TOP_AdaptabilityToMilking" "GBV_TOP_ShedTemperament"       "GBV_TOP_MilkingSpeed"         
# [64] "GBV_TOP_OverallOpinion"        "GBV_TOP_Stature"               "GBV_TOP_Capacity"             
# [67] "GBV_TOP_RumpAngle"             "GBV_TOP_RumpWidth"             "GBV_TOP_Legs"                 
# [70] "GBV_TOP_UdderSupport"          "GBV_TOP_ForeUdder"             "GBV_TOP_RearUdder"            
# [73] "GBV_TOP_FrontTeat"             "GBV_TOP_RearTeat"              "GBV_TOP_TeatLength"           
# [76] "GBV_TOP_UdderOverall"          "GBV_TOP_DairyConformation"     "PctHolstein"
```

# Diagnosis

## Herd size distribution in `r min(full_df$DairyYear)` to `r max(full_df$DairyYear)`

```{r herd size diagnosis}
for(i in min(full_df$DairyYear):max(full_df$DairyYear)) {
  herd_size <- full_df %>% 
  filter(DairyYear == i) %>% 
  group_by(herd_milk_type, HerdDurableKey) %>% 
  tally()

k <- ungroup(herd_size) %>% group_by(herd_milk_type) %>% 
  summarise(count = n(), min = min(n), mean = mean(n), sd = sd(n), max = max(n)) %>% 
  knitr::kable(digits = 0, 
               caption = paste0("Herd size by herd milking type in ", i)) %>% 
  kableExtra::kable_styling(full_width = T)
print(k)

p <- ggplot(herd_size, aes(n)) +
  geom_histogram(bins = 100) +
  # geom_vline(aes(xintercept = mean(n)), color = "red") +
  # geom_vline(aes(xintercept = mean(n)+2*sd(n)), color = "blue") +
  facet_grid(herd_milk_type~., scale = "free_y") + 
  labs(title = paste0("Herd size in ", i))
print(p)
p <- ggplot(herd_size, aes(n, fill = herd_milk_type)) +
  geom_boxplot() +
  labs(title = paste0("Herd size in ", i))
print(p)
}

  
```

## Holstein percentage distribution

```{r holstein perc}
holstein_pct <- full_df %>% 
  select(HerdDurableKey, AnimalDurableCode, PctHolstein, PctFriesian,
         herd_milk_type) %>% 
  distinct() %>% 
  group_by(herd_milk_type, HerdDurableKey) %>% 
  summarise(herd_foreign_pct = mean(PctHolstein), herd_hf_pct = mean(PctFriesian)
           )

ungroup(holstein_pct) %>% group_by(herd_milk_type) %>% 
  summarise(across(matches("herd_"), 
                   list(min = min, mean = mean, sd = sd, max = max))) %>% 
  knitr::kable(digits = 2, 
               caption = "Mean herd Holstein percentage by herd milking type") %>% 
  kableExtra::kable_styling(full_width = T)

for (i in c("herd_foreign_pct", "herd_hf_pct")) {
p <-  ggplot(holstein_pct, aes_string(i)) +
    geom_histogram(bins = 100) +
    facet_grid(herd_milk_type~., scale = "free_y") +
  # geom_vline(aes(xintercept = mean(mean)), color = "red") +
  # geom_vline(aes(xintercept = mean(mean)+2*sd(mean)), color = "blue") +
    labs(title = paste0("Mean ", i))
print(p)
p <- ggplot(holstein_pct, aes_string(i, fill = "herd_milk_type")) +
  geom_boxplot() +
  labs(title = paste0("Mean ", i))
print(p)
}

```

## Milk volume distribution

Cannot just extract Nov production because none of the months have records of
all cows.

```{r milk_volume}
test <- select(full_df, AnimalDurableCode, event_date) %>% 
  mutate(mo = lubridate::month(event_date))

n_animal <- unique(test$AnimalDurableCode) %>% length()

records <- data.frame(month = 1:12, n_cows = NA)
for (i in 1:12) {
  records$n_cows[i] <- filter(test, mo == i) %>% 
    select(AnimalDurableCode) %>% distinct() %>% nrow()
}
records$pct <- records$n_cows/n_animal
records %>% 
  knitr::kable(format.args = list(big.mark = ","), digits = 2, 
               caption = "N cows with milking records by month") %>% 
  kableExtra::kable_styling(full_width = T)

# Sept-Dec together
records <- data.frame(month = "9-12", n_cows = NA)
records$n_cows[1] <- filter(test, mo %in% 9:12) %>% 
    select(AnimalDurableCode) %>% distinct() %>% nrow()
records$pct <- records$n_cows/n_animal
records %>% 
  knitr::kable(format.args = list(big.mark = ","), digits = 2, 
               caption = "N cows with milking records in Setp-Dec") %>% 
  kableExtra::kable_styling(full_width = T)


```

### check Sept-Dec

For each cow, get the mean volume of all the herd tests in that month across
all years, i.e. life time mean volume by month.  
Then get the distribution of the above for each herd.

```{r milk_vol}
milk_vol <- select(full_df, HerdDurableKey, herd_milk_type, AnimalDurableCode,
                   event_date, Volume_ml) %>% 
  mutate(mo = lubridate::month(event_date)) 

test <- lapply(9:12, function(i) {
  out <- milk_vol %>% 
    filter(mo == i) %>% # can have records of the same mo in multiple years
    group_by(herd_milk_type, HerdDurableKey, AnimalDurableCode) %>% 
    summarise(vol = mean(Volume_ml))
  
  out$month <- i
  return(out)
})

milk_vol <- do.call(rbind, test)

ungroup(milk_vol) %>% group_by(month, herd_milk_type) %>% 
  summarise(min = min(vol), mean = mean(vol), sd = sd(vol),
            max = max(vol)) %>% 
  knitr::kable(format.args = list(big.mark = ","), digits = 0, 
               caption = "Mean cow milk volume by month and herd milking type") %>% 
  kableExtra::kable_styling(full_width = T)

# ggplot(milk_vol, aes(vol)) +
#   geom_density() +
#   facet_grid(month~herd_milk_type, scale = "free_y") +
#   # geom_vline(aes(xintercept = mean(mean)), color = "red") +
#   # geom_vline(aes(xintercept = mean(mean)+2*sd(mean)), color = "blue") +
#   labs(title = "Mean cow milk volumn")

ggplot(milk_vol, aes(vol, fill = herd_milk_type)) +
  geom_boxplot() + #orientation = ) +
  facet_grid(month~., scales = "free_y") +
  labs(title = "Mean cow milk volume")

```

# Data cleaning of TAD herds

```{r before_cleaning}
# sanity check
cat("Before cleaning.\n
    n animals:", length(unique(full_df$AnimalDurableCode)), ", n events:",
    nrow(full_df), ", n herds:\n")
group_by(full_df, herd_milk_type, HerdDurableKey) %>% 
  tally() %>% ungroup() %>% 
  group_by(herd_milk_type) %>% summarise(count = n()) %>% knitr::kable()
```

```{r remove herd size}
remove_herds <- lapply(min(full_df$DairyYear):max(full_df$DairyYear), 
                       function(i) {
  herd_size <- full_df %>% 
    filter(DairyYear == i) %>% 
    group_by(herd_milk_type, HerdDurableKey) %>% 
    tally()
  
  max_herd_size <- max(herd_size$n[herd_size$herd_milk_type=="OAD"])
  return(herd_size$HerdDurableKey[herd_size$n>max_herd_size &
                                  # herd_size$n<10 &
                                    herd_size$herd_milk_type=="TAD"])
                       })
remove_herds <- do.call(rbind, remove_herds)
  
df_new <- filter(full_df, !HerdDurableKey %in% remove_herds)
# sanity check
cat("Remove herds with size bigger than max OAD herd size in each year.
New data has:\n
    n animals:", length(unique(df_new$AnimalDurableCode)), ", n events:",
    nrow(df_new), ", n herds:\n")
group_by(df_new, herd_milk_type, HerdDurableKey) %>% 
  tally() %>% ungroup() %>% 
  group_by(herd_milk_type) %>% summarise(count = n()) %>% knitr::kable()
```

```{r remove_foreign_holstein}
max_foreign <- 0.35
df_new <- filter(df_new, PctHolstein < max_foreign)
# sanity check
cat("Remove herds with foreign Holstein % more than", max_foreign, ".
New data has:\n
    n animals:", length(unique(df_new$AnimalDurableCode)), ", n events:",
    nrow(df_new), ", n herds:\n")
group_by(df_new, herd_milk_type, HerdDurableKey) %>% 
  tally() %>% ungroup() %>% 
  group_by(herd_milk_type) %>% summarise(count = n()) %>% knitr::kable()
```

```{r remove holstein friesian}
max_hf <- 0.4
df_new <- filter(df_new, PctFriesian < max_hf)
# sanity check
cat("Remove herds with HF % more than", max_hf, ".
New data has:\n
    n animals:", length(unique(df_new$AnimalDurableCode)), ", n events:",
    nrow(df_new), ", n herds:\n")
group_by(df_new, herd_milk_type, HerdDurableKey) %>% 
  tally() %>% ungroup() %>% 
  group_by(herd_milk_type) %>% summarise(count = n()) %>% knitr::kable()
```

```{r remove high milk volume}
remove_herds <- lapply(9:12, function(i) {
  max_vol <- filter(milk_vol, month == i & herd_milk_type == "OAD") %>%
    ungroup() %>% 
    select(vol) %>% unlist() %>% max()
  
  return(filter(milk_vol, month == i & vol > max_vol) %>% 
           ungroup() %>% select(HerdDurableKey) %>% unlist())
})
remove_herds <- unlist(remove_herds) %>% unique()

df_new <- filter(df_new, !HerdDurableKey %in% remove_herds)
# sanity check
cat("Remove herds with milk volume more than max volume of OAD herds in each
month. New data has:\n
    n animals:", length(unique(df_new$AnimalDurableCode)), ", n events:",
    nrow(df_new), ", n herds:\n")
group_by(df_new, herd_milk_type, HerdDurableKey) %>% 
  tally() %>% ungroup() %>% 
  group_by(herd_milk_type) %>% summarise(count = n()) %>% 
  knitr::kable()
```
```{r save, eval = FALSE}
saveRDS(df_new, 
        paste0(fileDir,
               "production_survival_noYoungAnimal_gebv_beforeParity5_noOutliear.RData"))
```


```{r output, eval = FALSE}
rmarkdown::render(output_dir = outputDir, 
                  input = rstudioapi::getSourceEditorContext()$path)
```