---
title: "Summary statistics down sampled data"
author: "Luna Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
   # theme: cosmo
   # highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F,
                      root.dir = "e:/500390_OAD/output/sumstat_production/" )
```

```{r, include=F}
# Luna Zhang Feb 28 2022

# library(data.table)
suppressWarnings({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(ggpubr)
  library(tibble)
})

# source("utils.R")

fileDir <- "e:/500390_OAD/output/" # "z:/Luna/500390_OAD/explore/"
rawfileDir <- "z:/Luna/500390_OAD/OAD Dairy Data/"
outputDir <- "C:/Users/lzhang/AbacusBio Ltd/Pivot Cloud Drive - 500390 - DNZ TOPs impact on OAD survival 2021-3637/2. Doing/Results/" # "C:/Users/lzhang/Documents/GitHub/500390_OAD/output/"

if(!exists(outputDir)) {
  dir.create(outputDir)
}
# stop()
```

Read down-sampled dataset

```{r, include = F}
# df_downsample <- readRDS(
  # paste0(fileDir,
         # "production_survival_noYoungAnimal_gebv_beforeParity5_noOutliear.RData"))

df_downsample <- readRDS(paste0(fileDir, "modeling/df_downsample.RData"))
```


```{r, eval = F}
names(df_downsample)
#  [1] "AnimalDurableCode"             "DaysFromBirthToParturition"    "DaysFromParturitionToHerdTest"
#  [4] "InducedParturitionFlag"        "Volume_ml"                     "FatVolume_ml"                 
#  [7] "ProteinVolume_ml"              "AgeParity"                     "HerdDurableKey"               
# [10] "DairyYear"                     "Season"                        "EventDate"                    
# [13] "AgeParityGroup"                "RegimeType"                    "event_date"                   
# [16] "herd_milk_type"                "transition_year"               "s12"                          
# [19] "s23"                           "s34"                           "s45"                          
# [22] "EBV_BodyConditionScore"        "EBV_Fertility_CR42"            "EBV_Fertility_CSD0"           
# [25] "EBV_Fertility_CSD123"          "EBV_Fertility_PM123"           "EBV_FunctionalSurvival"       
# [28] "EBV_IndirectSurvival"          "EBV_SomaticCellScore"          "EBV_Survival"                 
# [31] "EBV_TOP_AdaptabilityToMilking" "EBV_TOP_Capacity"              "EBV_TOP_DairyConformation"    
# [34] "EBV_TOP_ForeUdder"             "EBV_TOP_FrontTeat"             "EBV_TOP_Legs"                 
# [37] "EBV_TOP_MilkingSpeed"          "EBV_TOP_OverallOpinion"        "EBV_TOP_RearTeat"             
# [40] "EBV_TOP_RearUdder"             "EBV_TOP_RumpAngle"             "EBV_TOP_RumpWidth"            
# [43] "EBV_TOP_ShedTemperament"       "EBV_TOP_Stature"               "EBV_TOP_TeatLength"           
# [46] "EBV_TOP_UdderOverall"          "EBV_TOP_UdderSupport"          "EBV_TotalFat"                 
# [49] "EBV_TotalProtein"              "EBV_TotalVolume"               "transition_parity"            
# [52] "EBV_Liveweight"                "GBV_TotalFat"                  "GBV_TotalProtein"             
# [55] "GBV_TotalVolume"               "GBV_Liveweight"                "GBV_SomaticCellScore"         
# [58] "GBV_Fertility_CR42"            "GBV_BodyConditionScore"        "GBV_FunctionalSurvival"       
# [61] "GBV_TOP_AdaptabilityToMilking" "GBV_TOP_ShedTemperament"       "GBV_TOP_MilkingSpeed"         
# [64] "GBV_TOP_OverallOpinion"        "GBV_TOP_Stature"               "GBV_TOP_Capacity"             
# [67] "GBV_TOP_RumpAngle"             "GBV_TOP_RumpWidth"             "GBV_TOP_Legs"                 
# [70] "GBV_TOP_UdderSupport"          "GBV_TOP_ForeUdder"             "GBV_TOP_RearUdder"            
# [73] "GBV_TOP_FrontTeat"             "GBV_TOP_RearTeat"              "GBV_TOP_TeatLength"           
# [76] "GBV_TOP_UdderOverall"          "GBV_TOP_DairyConformation"     "PctHolstein"        
# "PctFriesian" 
# [81] "EBV_TOP_AdaptabilityToMilking2" "EBV_TOP_Capacity2"             
 # [83] "EBV_TOP_DairyConformation2"     "EBV_TOP_ForeUdder2"            
 # [85] "EBV_TOP_FrontTeat2"             "EBV_TOP_Legs2"                 
 # [87] "EBV_TOP_MilkingSpeed2"          "EBV_TOP_OverallOpinion2"
 # [89] "EBV_TOP_RearTeat2"              "EBV_TOP_RearUdder2"
 # # [91] "EBV_TOP_RumpAngle2"             "EBV_TOP_RumpWidth2"            
 # [93] "EBV_TOP_ShedTemperament2"       "EBV_TOP_Stature2"              
 # [95] "EBV_TOP_TeatLength2"            "EBV_TOP_UdderOverall2"         
 # [97] "EBV_TOP_UdderSupport2"          "GBV_TOP_AdaptabilityToMilking2"
 # [99] "GBV_TOP_ShedTemperament2"       "GBV_TOP_MilkingSpeed2"         
# [101] "GBV_TOP_OverallOpinion2"        "GBV_TOP_Stature2"              
# [103] "GBV_TOP_Capacity2"              "GBV_TOP_RumpAngle2"            
# [105] "GBV_TOP_RumpWidth2"             "GBV_TOP_Legs2"                 
# [107] "GBV_TOP_UdderSupport2"          "GBV_TOP_ForeUdder2"            
# [109] "GBV_TOP_RearUdder2"             "GBV_TOP_FrontTeat2"            
# [111] "GBV_TOP_RearTeat2"              "GBV_TOP_TeatLength2"           
# [113] "GBV_TOP_UdderOverall2"          "GBV_TOP_DairyConformation2"    
```

# Holstein percentage distribution by milking type

```{r holstein}
holstein_pct <- df_downsample %>% 
  select(HerdDurableKey, AnimalDurableCode, PctHolstein, PctFriesian,
         herd_milk_type) %>% 
  distinct() %>% 
  mutate(herd_milk_type = factor(herd_milk_type, 
                                 levels = c("TAD", "OAD", "Transition"))) %>% 
  group_by(herd_milk_type, HerdDurableKey) %>% 
  summarise(herd_foreign_pct = mean(PctHolstein), herd_hf_pct = mean(PctFriesian)
           )

sumstat_holstein <- ungroup(holstein_pct) %>% group_by(herd_milk_type) %>% 
  summarise(across(matches("herd_"), 
                   list(min = min, mean = mean, sd = sd, max = max)))

write.csv(sumstat_holstein, paste0(outputDir, "sumstat_holstein.csv"), 
          row.names = F)

sumstat_holstein%>% 
  knitr::kable(digits = 2, 
               caption = "Mean herd Holstein percentage by herd milking type") %>% 
  kableExtra::kable_styling(full_width = T)

for (i in c("herd_foreign_pct", "herd_hf_pct")) {
# p <-  ggplot(holstein_pct, aes_string(i, fill = "herd_milk_type")) +
  # geom_boxplot() +
  # labs(title = paste0("Mean ", i)) +
  p <- ggboxplot(holstein_pct, "herd_milk_type", i, fill = "herd_milk_type", 
            title = paste0("Mean ", i), xlab = "", ggtheme = theme_pubclean()) +
    stat_compare_means(label = "p.signif", ref.group = "TAD",
                       method = "t.test") +
  scale_fill_manual(values = 
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(holstein_pct$herd_milk_type)-1, "Set3")))
  print(p)
}

```

# Milk production by milking type in Sept-Dec

For each cow, get the mean volume of all the herd tests in that month across
all years, i.e. life time mean volume by month.  
Then get the distribution of the above for each herd.

98% cows have herd tests in Sept, Oct, Nov or Dec, hence the 4 months.

```{r milk_vol}
milk_vol <- select(df_downsample, HerdDurableKey, herd_milk_type, AnimalDurableCode,
                   event_date, Volume_ml) %>% 
  mutate(mo = lubridate::month(event_date),
         herd_milk_type = factor(herd_milk_type, 
                                 levels = c("TAD", "OAD", "Transition"))) 

test <- lapply(9:12, function(i) {
  out <- milk_vol %>% 
    filter(mo == i) %>% # can have records of the same mo in multiple years
    group_by(herd_milk_type, HerdDurableKey, AnimalDurableCode) %>% 
    summarise(vol = mean(Volume_ml))
  
  out$month <- i
  return(out)
})

milk_vol <- do.call(rbind, test)

sumstat_milk_vol <- ungroup(milk_vol) %>% group_by(month, herd_milk_type) %>% 
  summarise(min = min(vol), mean = mean(vol), sd = sd(vol),
            max = max(vol)) 

sumstat_milk_vol%>% 
  knitr::kable(format.args = list(big.mark = ","), digits = 0, 
               caption = "Mean cow milk volume by month and herd milking type") %>% 
  kableExtra::kable_styling(full_width = T)

# ggplot(milk_vol, aes(vol)) +
#   geom_density() +
#   facet_grid(month~herd_milk_type, scale = "free_y") +
#   # geom_vline(aes(xintercept = mean(mean)), color = "red") +
#   # geom_vline(aes(xintercept = mean(mean)+2*sd(mean)), color = "blue") +
#   labs(title = "Mean cow milk volumn")

ggboxplot(milk_vol, "herd_milk_type", "vol", fill = "herd_milk_type", 
            title = "Mean cow milk volume", xlab = "", ylim = c(0, 50000),
          facet.by = "month", ggtheme = theme_pubclean()) +
    stat_compare_means(aes(label = ..p.signif..), ref.group = "TAD",
                       method = "t.test") +
  scale_fill_manual(values = 
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(holstein_pct$herd_milk_type)-1, "Set3")))


```

```{r milk vol perc}
sumstat_milk_vol$percent <- 0
for (i in min(sumstat_milk_vol$month):max(sumstat_milk_vol$month)) {
  sumstat_milk_vol$percent[sumstat_milk_vol$month==i] <- 
    sumstat_milk_vol$mean[sumstat_milk_vol$month==i]/
    sumstat_milk_vol$mean[sumstat_milk_vol$herd_milk_type=="TAD" &
                            sumstat_milk_vol$month==i]*100
}
write.csv(sumstat_milk_vol, paste0(outputDir, "sumstat_milk_vol.csv"),
          row.names = F)

sumstat_milk_vol$percent <- round(sumstat_milk_vol$percent, 2)

ggbarplot(sumstat_milk_vol, "month", "percent", fill = "herd_milk_type", 
          color = "black", 
          label = rep(c("", "****", "****"), 4), # TRUE, 
          position = position_dodge(0.9),
          title = "Percent difference in monthly mean milk volume") +
  scale_fill_manual(values = 
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(sumstat_milk_vol$herd_milk_type)-1, "Set3"))) 
```

# Herd size distribution by milking type

## N herds

```{r herd n}
size <- df_downsample %>% group_by(DairyYear, herd_milk_type, HerdDurableKey) %>% 
  summarise(count = n()) %>% ungroup() %>% 
  group_by(DairyYear, herd_milk_type) %>% 
  tally() %>% ungroup() %>% 
  group_by(herd_milk_type) %>% 
  summarise(count = n(), min = min(n), median = median(n), max = max(n), 
            mean = mean(n), SD = sd(n))
write.csv(size, paste0(outputDir, "herd_size_by_year.csv"), row.names = F)

size <- df_downsample %>% group_by(DairyYear, herd_milk_type, HerdDurableKey) %>% 
  summarise(count = n()) %>% ungroup() %>% 
  group_by(DairyYear, herd_milk_type) %>% 
  tally() %>% 
  mutate(herd_milk_type = factor(herd_milk_type, 
                                 levels = c("Transition", "OAD", "TAD")),
         DairyYear = factor(DairyYear))

ggline(size, "DairyYear", "n", add = "mean_se", color = "herd_milk_type",
          palette = "jco",
          xlab = "")# +
  # stat_compare_means(aes(group = herd_milk_type), label = "p.signif", method = "t.test",
                     # label.y = 450) + # rep(500, nlevels(size$DairyYear))) 
```

## mean across years

```{r herd size}
group_by(df_downsample, new_milk_type, DairyYear) %>% 
 # summarise(count = n()) %>% 
  #ungroup() %>% group_by(herd_milk_type, new_milk_type) %>% 
  tally() %>% ungroup() %>% 
  group_by(new_milk_type) %>% 
  summarise(count = round(mean(n), digits = 0)) %>% ungroup() %>% 
  mutate(prop = count/sum(count)) %>% 
  knitr::kable(align = "l",
               caption = "Mean N animals by milking type across years", 
               digits = 3,
               format.args = list(big.mark=",")) %>% 
  kable_styling(full_width = T)
```

## per year

```{r herd size1, cache = T}
for(i in min(df_downsample$DairyYear):max(df_downsample$DairyYear)) {
  herd_size <- df_downsample %>% 
    filter(DairyYear == i) %>% 
    group_by(herd_milk_type, HerdDurableKey) %>% 
    tally() %>% 
    mutate(mean = mean(n), median = median(n))

# ungroup(herd_size) %>% group_by(herd_milk_type) %>% 
#     summarise(count = n(), min = min(n), mean = mean(n), sd = sd(n),
#               max = max(n)) %>%
#     knitr::kable(digits = 0, 
#                  caption = paste0("Mean herd size by herd milking type in ", i)) %>% 
#     kableExtra::kable_styling(full_width = T)

p <- ggplot(herd_size, aes(n)) +
  geom_histogram(bins = 100) +
  facet_grid(herd_milk_type~., scale = "free_y") +
  geom_vline(aes(xintercept = mean, group = herd_milk_type), color = "red") +
  geom_vline(aes(xintercept = median, group = herd_milk_type), color = "blue") +
  # geom_vline(aes(xintercept = mean(n)+2*sd(n)), color = "blue") +
  labs(title = i)

# p <- herd_size %>% 
  # gghistogram(x = "n", binwidth = 5, add = "mean", fill = "herd_milk_type",
  #             title = i
  #             ) %>% 
  # facet("herd_milk_type", nrow = 3, ncol = 1, scales = "free_y") %>% 
  # ggpar(legend = "none")

  print(p)
}
```

# survival rate vs milking type

*Animals with missing survival are removed.*

```{r survival, cache = T, echo = F}
sur_milk_type <- lapply(2:6, function(i) {
  out <- df_downsample %>% filter(AgeParity==i) %>% 
    filter(across(paste0("s", i-1, i), ~.%in%c(T, F))) %>% 
    group_by(new_milk_type, across(paste0("s", i-1, i))) %>% 
    tally() %>% 
    mutate(prp = round(n/sum(n), digits = 2),
           survival = paste0("s", i-1, i))
  names(out)[2] <- "survived"
  
  return(out)
})
sur_milk_type <- do.call(rbind, sur_milk_type)

# rate
test <- filter(sur_milk_type, survived == T) %>% 
  select(-survived, -n) %>% 
  pivot_wider(c("survival"), "new_milk_type", values_from = "prp" ) %>% 
  select(survival, TAD, everything())
knitr::kable(test) %>% 
  kableExtra::kable_styling()
write.csv(test, paste0(outputDir, "survival_rate_in_milke_type.csv"))

# Proportion test
sur_milk_type_n <- sur_milk_type %>% 
  pivot_wider(c("survival", "new_milk_type"), "survived", values_from = "n" )
sur_milk_type_n[["FALSE"]][which(is.na(sur_milk_type_n[["FALSE"]]))] <- 0
sur_milk_type_n$total <- sur_milk_type_n$`FALSE`+sur_milk_type_n$`TRUE`
sur_milk_type_n$prp <- sur_milk_type_n$`TRUE`/sur_milk_type_n$total
sur_milk_type_n <- select(sur_milk_type_n, -`FALSE`)

prop <- lapply(split(sur_milk_type_n, sur_milk_type_n$survival), function(i) {
  
  sub_prop <- lapply(setdiff(i$new_milk_type, "TAD"), function(j) {
    freq <- filter(i, new_milk_type %in% c("TAD", j)) %>% 
      ungroup() %>% 
      select(`TRUE`, total)
    return(prop.test(freq$`TRUE`, freq$total)$p.value)
  })
  
  p_value <- data.frame(new_milk_type = setdiff(i$new_milk_type, "TAD"),
                        p_value = unlist(sub_prop))
  
  return(left_join(i, p_value))
})
prop <- do.call(rbind, prop)

```

## TAD and OAD only

```{r survival bar, fig.show='asis'}
df <- prop %>% 
  mutate(sign = ifelse(p_value < 0.001, "***",
                         ifelse(p_value < 0.01, "**",
                                ifelse(p_value < 0.05, "*", 
                                       ifelse(p_value < 0.06, ".", "")))),
           new_milk_type = 
             factor(new_milk_type, 
                    levels = c("TAD", setdiff(unique(prop$new_milk_type), "TAD"))))

ggplot(filter(df, as.integer(new_milk_type) <=2),
       aes(survival, prp, fill = new_milk_type)) +
  geom_bar(stat = "identity", width = .8, color = "black", position = position_dodge(.8)) +
  geom_text(aes(label = sign), position = position_dodge(.8),
            vjust = 0) + #-0.05) +
  theme_pubclean() +
  scale_fill_manual(values = 
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(df$new_milk_type)-1, "Set3")))

```

## all milking types

```{r survival_bar1, fig.show='asis'}
p <- ggplot(df, aes(survival, prp, fill = new_milk_type)) +
  geom_bar(stat = "identity", width = .8, color = "black", position = position_dodge(.8)) +
  geom_text(aes(label = sign), position = position_dodge(.8),
            vjust = 0) + #-0.05) +
  theme_pubclean() +
  scale_fill_manual(values = 
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(df$new_milk_type)-1, "Set3")))
print(p)
```

## table

```{r}
names(prop)[3] <- "survived"
knitr::kable(prop, digits =2, format.args = list(big.mark = ",")) %>% 
  kableExtra::kable_styling()

write.csv(prop, paste0(outputDir, "survival_in_milk_type.csv"),
          quote = F, row.names = F)
```

# EBV distribution within new_milk_type

```{r ebv, cache=TRUE}
sumstat_ebv <- lapply(grep("EBV_", names(df_downsample), value = T), function(i) {
  
  # table for plot
  sumstat_ebv <- select(df_downsample, ends_with(i), # starts_with("EBV_"),
                      new_milk_type) %>% #, new_milk_type) %>% 
    group_by(new_milk_type) %>% 
    summarise(across(everything(), 
                     list(mean = mean, median = median, SD = sd))) %>% 
    pivot_longer(-new_milk_type, "stat", paste0(i, "_"), values_to = i) 

  # table to print
  out <- sumstat_ebv %>%
    pivot_wider(stat, new_milk_type, values_from = matches(i)) %>% 
    mutate(trait = i ) 
  
  # # calculate t test p value
  # ps <- lapply(grep("OAD$|Tran", names(out), value = T), function(j) {
  #   return(t.test(as.formula(paste0(i, "~new_milk_type")),
  #          data = select(df_downsample, new_milk_type, matches(i)) %>% 
  #            filter(new_milk_type %in% c("TAD", j)))$p.value)
  # })
  # 
  # # add p to both tables
  # p_value <- data.frame(new_milk_type = grep("OAD$|Tran", names(out), value = T),
  #                  stat = rep("p_value", 
  #                             grepl("OAD$|Tran", names(out)) %>% sum()),
  #                  value = unlist(ps), 
  #                  trait = i)
  names(sumstat_ebv)[3] <- "value"
  sumstat_ebv$trait <- i
  # sumstat_ebv <- rbind(sumstat_ebv, p_value)
  # 
  # idx <- grep("OAD$|Tran", names(out))
  # p_value <- out[1,]
  # for(i in seq(idx)) p_value[1,idx[i]] <- ps[[i]]
  # p_value[1,1] <- "p_value"
  # p_value$TAD[1] <- NaN
  # 
  # out <- rbind(out, p_value)
  # 
 return(list(for_plot = sumstat_ebv, out = out))
})

ebv_table <- do.call(rbind, lapply(sumstat_ebv, `[[`, "out")) %>% 
  select(-matches("NA")) %>% 
  select(trait, stat, TAD, everything())

write.csv(ebv_table, paste0(outputDir, "sumstat_ebv.csv"), quote = F, row.names = F)

# print kable
knitr::kable(
  x = ebv_table, digits = 2,
  align = "l", format.args = list(big.mark = ","),
  caption = "") %>%
  kableExtra::kable_styling(fixed_thead = T)

```

# GBV distribution within new_milk_type

```{r ebv1, cache=TRUE}
sumstat_gbv <- lapply(grep("GBV_", names(df_downsample), value = T), function(i) {
  
  # table for plot
  sumstat_gbv <- select(df_downsample, ends_with(i), # starts_with("EBV_"),
                      new_milk_type) %>% #, new_milk_type) %>% 
    group_by(new_milk_type) %>% 
    summarise(across(everything(), 
                     list(mean = mean, median = median, SD = sd))) %>% 
    pivot_longer(-new_milk_type, "stat", paste0(i, "_"), values_to = i) 

  # table to print
  out <- sumstat_gbv %>%
    pivot_wider(stat, new_milk_type, values_from = matches(i)) %>% 
    mutate(trait = i ) 
  # 
  # # calculate t test p value
  # ps <- lapply(grep("OAD$|Tran", names(out), value = T), function(j) {
  #   return(t.test(as.formula(paste0(i, "~new_milk_type")),
  #          data = select(df_downsample, new_milk_type, matches(i)) %>% 
  #            filter(new_milk_type %in% c("TAD", j)))$p.value)
  # })
  # 
  # # add p to both tables
  # p_value <- data.frame(new_milk_type = grep("OAD$|Tran", names(out), value = T),
  #                  stat = rep("p_value", 
  #                             grepl("OAD$|Tran", names(out)) %>% sum()),
  #                  value = unlist(ps), 
  #                  trait = i)
  names(sumstat_gbv)[3] <- "value"
  sumstat_gbv$trait <- i
  # sumstat_gbv <- rbind(sumstat_gbv, p_value)
  # 
  # idx <- grep("OAD$|Tran", names(out))
  # p_value <- out[1,]
  # for(i in seq(idx)) p_value[1,idx[i]] <- ps[[i]]
  # p_value[1,1] <- "p_value"
  # p_value$TAD[1] <- NaN
  # 
  # out <- rbind(out, p_value)
  # 
 return(list(for_plot = sumstat_gbv, out = out))
})

gbv_table <- do.call(rbind, lapply(sumstat_gbv, `[[`, "out")) %>% 
  select(-matches("NA")) %>% 
  select(trait, stat, TAD, everything())

write.csv(gbv_table, paste0(outputDir, "sumstat_gbv.csv"), quote = F, row.names = F)

# print kable
knitr::kable(
  x = gbv_table, digits = 2,
  align = "l", format.args = list(big.mark = ","),
  caption = "") %>%
  kableExtra::kable_styling(fixed_thead = T)

```

# EBV vs GBV comparision

## Correlation

The reference line is $y=x$.

```{r bv_correlation, fig.show='asis', fig.height=12, fig.width=8, cache=TRUE}
traits <- c("BodyConditionScore", "Fertility_CR42", "FunctionalSurvival", 
            "SomaticCellScore", "TotalVolume", "TotalFat", "TotalProtein", 
            "TOP_UdderSupport", "TOP_MilkingSpeed", "TOP_Capacity", "TOP_FrontTeat")

# very slow
# probably because too many obs

df <- select(df_downsample, AnimalDurableCode, matches("BV_")) %>% 
  distinct()

ps <- lapply(traits, function(trait) {
  return(select(df, matches(paste0(trait, "$"))) %>% 
    ggscatter(paste0("EBV_", trait), paste0("GBV_", trait), 
              color = "black", shape = 21, size = 1, # Points color, shape and size
             # add = "reg.line",
            #  add.params = list(color = "aquamarine3", fill = "lightgray"), # Customize reg. line
            #  conf.int = TRUE, # Add confidence interval
              cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
              cor.coeff.args = list(method = "pearson", label.sep = "\n"),
              title = trait, xlab = "EBV", ylab = "GEBV") + 
      geom_abline(intercept = 0, slope = 1, color = "aquamarine4", size = 1.5))
})

invisible(gridExtra::marrangeGrob(grobs = ps, nrow = 4, ncol = 3) %>% print())

```

## Box plot

It takes too long to plot so only plot one trait

```{r ebv_boxplot, cache = TRUE, eval = F}
# Error in lazyLoadDBinsertValue(data, datafile, ascii, compress, envhook) : 
  # long vectors not supported yet: connections.c:6011
i <- grep("EBV_", names(df_downsample), value = T)[1]
  p <- ggboxplot(filter(df_downsample, !is.na(new_milk_type)),
                 "new_milk_type", i, fill = "new_milk_type",
                 palette = "jco") +
    stat_compare_means(method = "anova") +
    stat_compare_means(label = "p.signif", method = "t.test", ref.group = "TAD")
  print(p)

```

## comparison of mean EBV and GEBV across milking types

error bar chart 

```{r EBV_barchart}

ebv_plot <-  do.call(rbind, lapply(sumstat_ebv, `[[`, "for_plot")) 
gbv_plot <-  do.call(rbind, lapply(sumstat_gbv, `[[`, "for_plot")) 
ebv_plot$evaluation <- "EBV"
gbv_plot$evaluation <- "GEBV"

for_plot <- rbind(ebv_plot, gbv_plot)
for_plot$trait <- strsplit(for_plot$trait, "BV_") %>% sapply(tail, 1)

traits <- grep("BV_", names(df_downsample), value = T) %>% 
  strsplit("BV_") %>% sapply(tail, 1) %>% unique()

if(length(unique(for_plot$trait))!=length(traits)) stop("EBV_barchart, trait n inconsistent")
# t test GBV vs EBV
p_values <- lapply(unique(traits), function(i) {
  
  df <- df_downsample %>% 
    select(new_milk_type, AnimalDurableCode, matches(paste0("BV_", i, "$"))) %>% 
    distinct() %>% 
    pivot_longer(matches(i), "evaluation", paste0("_", i), values_to = "value")
  
  if(df$evaluation %>% unique() %>% length()<2) {return(NULL)}
  
  by_milk <- lapply(unique(df$new_milk_type), function(j) {
    df_sub <- filter(df, new_milk_type == j)
    return(t.test(as.formula(value~evaluation), data = df_sub)$p.value)
  })
  return(data.frame(trait = rep(i,length(by_milk)),
                    new_milk_type = unique(df$new_milk_type),
                    p_value = unlist(by_milk)))
})
p_values <- do.call(rbind.data.frame, p_values)

ps <- lapply(unique(for_plot$trait), function(i) {
  df <- filter(for_plot, trait == i) %>% 
    pivot_wider(c("trait", "new_milk_type", "evaluation"), stat,
                values_from = "value") %>% 
    left_join(p_values) %>% 
    mutate(sign = ifelse(p_value < 0.001, "***",
                         ifelse(p_value < 0.01, "**",
                                ifelse(p_value < 0.05, "*", ""))),
           new_milk_type = 
             factor(new_milk_type, 
                    levels = c("TAD", setdiff(unique(new_milk_type), "TAD"))))
  
  df$sign[df$evaluation=="EBV"] <- ""
  
  return(list(df = df,
              p = ggplot(df, aes(new_milk_type, mean, color = evaluation)) +
                geom_pointrange(aes(ymin = mean-SD, ymax = mean+SD),
                                position = position_dodge(0.5)) +
                # geom_errorbar(aes(ymin = mean-SD, ymax = mean+SD), width = .2) +
                labs(title = i, x = "") +
                geom_text(aes(y = mean+SD, label = sign),
                          position = position_dodge(1), vjust = 0) + #-0.5) +
                theme_classic()))
})

invisible(sapply(lapply(ps, `[[`, "p"), print))

sumstat_gbv_ebv <- do.call(rbind.data.frame, lapply(ps, `[[`, "df"))
write.csv(sumstat_gbv_ebv, paste0(outputDir, "ebv_gbv_compare.csv"))

```

# Compare mean EBV of survived and dead cows 

```{r ebv within sur and milk, results='asis'}

ebv_sur_milk <- lapply(2:5, function(i) {
  out <- df_downsample %>% filter(AgeParity==i) %>% 
    filter(across(paste0("s", i-1, i), ~.%in%c(T, F))) %>% 
    select(new_milk_type, starts_with("s"), starts_with("EBV")) %>% 
    group_by(new_milk_type, across(paste0("s", i-1, i))) %>% 
    summarise(across(starts_with("EBV"), ~mean(.), na.rm = T)) %>% 
    mutate(survival = paste0("s", i-1, i))
  names(out)[2] <- "survived"
  
  return(out)
})
ebv_sur_milk <- do.call(rbind, ebv_sur_milk)

longer <- ebv_sur_milk %>% 
  pivot_longer(starts_with("EBV"), "trait", values_to = "mean_EBV") %>% 
  arrange(survival, new_milk_type, trait, survived) %>% 
  pivot_wider(names_from = survived, values_from = mean_EBV, 
              names_prefix = "survived_")

# longer$diff <- longer$survived_FALSE-longer$survived_TRUE
# knitr::kable(longer, digits = 3)
```

```{r t test, cache=TRUE}
by_sur <- lapply(2:5, function(i) {
  
  by_trait <- lapply(grep("EBV_", names(df_downsample), value = T), function(j) {
    
    df <- filter(df_downsample, AgeParity==i) %>% 
      select(matches(paste0("s", i-1, i)), new_milk_type, matches(j),
             AnimalDurableCode) %>% 
      pivot_longer(starts_with("s"), "survival", values_to = "value") %>% 
      distinct() 
    
    by_milk <- lapply(unique(df$new_milk_type), function(k) {
      df_sub <- filter(df, new_milk_type == k)
      
      if(T %in% df_sub$value && F %in% df_sub$value) {
        return(
          data.frame(trait = j,
                     new_milk_type = k,
                     p_value = t.test(as.formula(paste0(j,"~value")), 
                                      data = filter(df_sub, !is.na(value)))$p.value))
      } else {
        return(NULL)
      }
    })
    out <- do.call(rbind.data.frame, by_milk)
    
    if(length(out) > 0) {
      return(out %>% mutate(survival = paste0("s", i-1, i)))
    }
  })
  return(do.call(rbind.data.frame, by_trait))
})
by_sur <- do.call(rbind.data.frame, by_sur)

longer <- left_join(longer, by_sur)
write.csv(longer, paste0(outputDir, "EBV_in_survival_milk_type.csv"),
          quote = F, row.names = F)

```

```{r ebv plot1, fig.show='asis', cache=TRUE}
df <- longer %>% 
  mutate(sign = ifelse(p_value < 0.001, "***",
                         ifelse(p_value < 0.01, "**",
                                ifelse(p_value < 0.05, "*", "")))) %>% 
  pivot_longer(cols = starts_with("survived_"), names_to = "survived",
               names_prefix = "survived_",
               values_to = "value")
df$sign[df$survived==F] <- ""

ps <- lapply(grep("EBV_", unique(df$trait), value = T), function(i) {
  df_sub <- filter(df, trait == i)
  mutate(df_sub, new_milk_type = 
             factor(new_milk_type, 
                    levels = c("TAD", setdiff(unique(df_sub$new_milk_type), "TAD")))) %>% 
  ggplot(aes(new_milk_type, value, fill = survived)) +
  geom_bar(stat = "identity", width = .8, color = "black", 
           position = position_dodge(.8)) +
  geom_text(aes(label = sign), position = position_dodge(.8),
            vjust = +0.5) +
  facet_grid(survival~.) +
  labs(title = i, x = "") +
  theme_pubclean() +
  scale_fill_manual(values = 
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(prop$new_milk_type)-1, "Set3")))
})
sapply(ps, print)
```


```{r, eval = FALSE}
rmarkdown::render(output_dir = outputDir, 
                  input = rstudioapi::getSourceEditorContext()$path)
```
