---
title: "Modelling"
author: "Luna Zhang"
date: "`r Sys.Date()`"
output: #> rmarkdown::html_vignette
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F, warning = F, message = F,
                      root.dir = "e:/500390_OAD/output/modeling/"
)
```

Modelling of survival on effects.

```{r, include =FALSE}
library(dplyr)
library(ggplot2)
library(plm)

source("utils.R")

fileDir <- "e:/500390_OAD/output/" # "z:/Luna/500390_OAD/explore/"
# rawfileDir <- "e:/OAD Dairy Data/2022021110_EBVs/"
outputDir <- "e:/500390_OAD/output/modeling/" # copy to "z:/Luna/500390_OAD/output/" later

# if(!exists(outputDir)) dir.create(outputDir)

memory.limit()/1024
memory.limit(size = 191/2*1024^2) # cannot be too big. Will crash
options(max.print = 100)
```

```{r read_RData, cache=TRUE, cache.lazy=FALSE, cache.extra=tools::md5sum("e:/500390_OAD/output/production_survival_noYoungAnimal_ebv_beforeParity5_n.RData")}
full_df <- readRDS(
  paste0(fileDir, "production_survival_noYoungAnimal_gebv_beforeParity5_noOutliear.RData"))

```

```{r sanity check}
names(full_df)
```

## make nonlinear traits

```{r nonlinear, cache = TRUE}
t <- Sys.time()
for(i in grep("EBV_TOP_", names(full_df), value = T)) {
  full_df[,paste0(i, "2")] <- full_df[[i]]^2+full_df[[i]]
}
print(Sys.time()-t)

for(i in grep("GBV_TOP_", names(full_df), value = T)) {
  full_df[,paste0(i, "2")] <- full_df[[i]]^2+full_df[[i]]
}

print(grep("BV", names(full_df), value = T))
```

## Down sample due to too many herds that uses up all memories

Down sample total herd size to 1000, otherwise will use up all the memories.
TAD herd size to 296. See `0062_sumstat.rmd`

```{r down_sample, eval = FALSE, cache = TRUE}
# set.seed(123)
# idx <- lapply(unique(full_df$herd_milk_type), function(i) { # 3 types
#   print(i)
#   idx <- which(full_df$herd_milk_type==i)
#   herd_ <- full_df$HerdDurableKey[idx] %>% unique()
#   herd_sample <- sample(herd_, min(333, length(herd_)), replace = F)
#   print(length(herd_sample))
#   idx <- which(full_df$HerdDurableKey %in% herd_sample)
#   return(idx)
# })
# # sanity check
# for(i in sample(1:3, 3, F)) {
#   for(j in sample(setdiff(1:3, i), 2)) {
#     match(idx[[i]], idx[[j]]) %>% unique() %>% print()   
#   }
# }
# idx <- unlist(idx)
# df_downsample <- full_df[sort(idx),]
set.seed(123)
idx_leftover <- which(full_df$new_milk_type!="TAD")
idx_tad <- which(full_df$new_milk_type=="TAD")
herd_tad <- full_df$HerdDurableKey[idx_tad] %>% unique()
herd_sample <- sample(herd_tad, 
                      full_df$HerdDurableKey[which(full_df$new_milk_type=="OAD")] %>% 
                        unique() %>% length(), replace = F)
idx <- which(full_df$HerdDurableKey %in% herd_sample)

# sanity check
match(idx, idx_leftover) %>% unique() # NA

df_downsample <- full_df[sort(c(idx_leftover, idx)),]

saveRDS(df_downsample, paste0(outputDir, "df_downsample.RData"))
```

```{r read downsample file}
df_downsample <- readRDS(paste0(outputDir, "df_downsample.RData"))
```

```{r downsample statistics}
rm(list=ls(pattern = "full_df"))
cat("Downsampled dataset has", nrow(df_downsample), "events,",
    unique(df_downsample$AnimalDurableCode) %>% length(), "cows, and",
    unique(df_downsample$HerdDurableKey) %>% length(), "herds.")

table(df_downsample$new_milk_type)

group_by(df_downsample, new_milk_type, HerdDurableKey) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% group_by(new_milk_type) %>% tally()

df_downsample <- mutate(df_downsample, 
                        cg = paste0(HerdDurableKey, "_", DairyYear))
```
## Model

### Two model distribution choices
* Logistic model, since survival is a binary trait.i.e.
log odds: $ln(odds) = ln(\frac{p}{1-p}) = a*x_1 + b*x_2 + … + z*x_n$  

The problem is it uses more memories. The cg level is too much and can cause
crash.

* Gaussian model, don't result in crash but doesn't assume variation in the
relationship between the effects and the survival (because it doesn't treat
survival as a continuous probability)

```{r distribution choice, include = TRUE}
distribution <- "gaussian" # logistic
```

### data points

Because each cow has multiple testing record, but their survival doesn't 
necessarily need call testing record, the data is reduced to one data point per
cow.

For s12-s45, filter down to AgeParity==2, 3, 4, or 5. Because at AgeParity==3,
all existing cows' s12 are 1s.

## Survival s12

```{r s12_filter, warning =FALSE, message =FALSE}
df_model <- select(df_downsample, AnimalDurableCode, HerdDurableKey, 
                   DairyYear, AgeParity, cg, 
                   new_milk_type, s12, starts_with("EBV_")) %>% 
  filter(AgeParity==2) %>% 
  distinct() %>% 
  mutate(HerdDurableKey = as.factor(HerdDurableKey), 
         DairyYear = as.factor(DairyYear))

which(duplicated(df_model$AnimalDurableCode)) # sanity check

cat("s12 dataset has", nrow(df_model), "events,",
    unique(df_model$AnimalDurableCode) %>% length(), "cows, and",
    unique(df_model$HerdDurableKey) %>% length(), "herds.")
```

### First, use all the cows to check if milking (regime type) has an effect

The first factor is CG.
can't fit Herdurable Key. use more than 500 GB RAM.
So, fit herd and dairy year separately.
DairyYear 2021 doesn't have survival data so it is excluded from the model.

Below is to confirm that Herd and DairyYear has an effect on s12.
Some herds do not have an effect but all DairyYears have effects.

```{r s12_cg, cache = TRUE}
# https://www.r-bloggers.com/2015/09/how-to-perform-a-logistic-regression-in-r/
# glm_s12_on_cg <- glm(s12~HerdDurableKey + DairyYear + 0, 
                     # binomial(link = "logit"), df_model)
# summary(glm_s12_on_cg)
# aov(glm_s12_on_cg) %>% summary()

s <- glm(formula = ifelse(distribution=="gaussian", s12~cg + 0, 
                          DairyYear + 0), 
         family = ifelse(distribution=="gaussian", gaussian(link = "identity"), 
                         binomial(link = "logit")), 
         df_model)
summary(s)
anovas <- anova(s, test = "Chisq")
print(anovas)

s <- glm(formula = ifelse(distribution=="gaussian", s12~cg + 0, 
                          DairyYear + 0), 
         family = ifelse(distribution=="gaussian", gaussian(link = "identity"), 
                         binomial(link = "logit")), 
         df_model)
summary(s)
anovas <- anova(s, test = "Chisq")
print(anovas)

```
### add milking type to the model

Fitting CG as combination of herd and year takes too long. Therefore has to
fit them separately.

To show milking type effects fully, fit it before herd effect in the model.

```{r s12_milk_type, cache=TRUE, eval = FALSE}
glm_s12_on_year_milktype <- glm(formula = ifelse(distribution=="gaussian", 
                                                 s12~cg+new_milk_type + 0,s12~DairyYear+new_milk_type+HerdDurableKey+ 0, 
# add 2007 into the equation by +0, to make the intercept == OAD
                              gaussian(link = "identity"), #  binomial(link = "logit"), 
df_model)
s <- summary(glm_s12_on_year_milktype)
print(s$call)
# cat("deviance residuals:\n"); summary(s$deviance.resid)
coefs <- data.frame(s$coefficients[grepl("new_milk_type", rownames(s$coefficients)),])
names(coefs)[4] <- "p-value"
coefs$sign <- ifelse(coefs$`p-value`<0.001, "***", 
                     ifelse(coefs$`p-value`<0.01, "**",
                            ifelse(coefs$`p-value`<0.05, "*",
                                   ifelse(coefs$`p-value`<0.1, ".",""))))
knitr::kable(coefs)
cat("\nAIC:",s$aic)
cat("\n\n##--ANOVA with Chi-sq test--##\n\n")
anovas <- anova(glm_s12_on_year_milktype, test = "Chisq")
print(anovas)

coefs$survival <- "s12"; anovas$survival <- "s12"
write.csv(coefs, paste0(outputDir, "coef_s12_milk_type.csv"), 
          row.names = T, quote = F)
write.csv(coefs, paste0(outputDir, "anova_s12_milk_type.csv"), 
          row.names = T, quote = F)
```

The interception is OAD and HerdDurableKey==145. 
This is to say when all other effects are at the same level, TAD decreases the 
odd ratio of S12 1:0 by around 2, meaning they increase the change of mortality.
Other milking types (transitioning) increases the odd ratio but their effects 
aren't significant. 

## S23

```{r s23_milk_type, cache = TRUE, eval = FALSE}
rm(list=ls(pattern = "idx_"))
df_model <- select(df_downsample, AnimalDurableCode, HerdDurableKey, 
                   DairyYear, AgeParity,
                   new_milk_type, s23, starts_with("EBV_")) %>% 
  filter(AgeParity==3 & !is.na(s23)) %>% 
  distinct() %>% 
  mutate(HerdDurableKey = as.factor(HerdDurableKey), 
         DairyYear = as.factor(DairyYear))

# which(duplicated(df_model$AnimalDurableCode)) # sanity check
cat("s23 dataset has", nrow(df_model), "events,",
    unique(df_model$AnimalDurableCode) %>% length(), "cows, and",
    unique(df_model$HerdDurableKey) %>% length(), "herds.")

glm_s23_on_year_milktype <- glm(s23~DairyYear+new_milk_type+HerdDurableKey+ 0, 
# add 2007 into the equation by +0, to make the intercept == OAD
                                binomial(link = "logit"), df_model)
s <- summary(glm_s23_on_year_milktype)
print(s$call)
# cat("deviance residuals:\n"); summary(s$deviance.resid)
coefs <- data.frame(s$coefficients[grepl("new_milk_type", rownames(s$coefficients)),])
coefs$sign <- ifelse(coefs$Pr...z..<0.001, "***", 
                     ifelse(coefs$Pr...z..<0.01, "**",
                            ifelse(coefs$Pr...z..<0.05, "*",
                                   ifelse(coefs$Pr...z..<0.1, ".",""))))
knitr::kable(coefs)
cat("\nAIC:",s$aic)
cat("\n##--ANOVA with Chi-sq test--##\n\n")
anovas <- anova(glm_s23_on_year_milktype, test = "Chisq")
print(anovas)

coefs$survival <- "s23"; anovas$survival <- "s23"
write.csv(coefs, paste0(outputDir, "coef_s23_milk_type.csv"), 
          row.names = T, quote = F)
write.csv(coefs, paste0(outputDir, "anova_s23_milk_type.csv"), 
          row.names = T, quote = F)
```

## S34

milking type has no effect on s34

```{r s34_milk_type, cache=TRUE, eval = FALSE}
df_model <- select(df_downsample, AnimalDurableCode, HerdDurableKey, 
                   DairyYear, AgeParity,
                   new_milk_type, s34, starts_with("EBV_")) %>% 
  filter(AgeParity==4) %>% 
  distinct() %>% 
  # mutate(cg = paste0(HerdDurableKey, "_", DairyYear))
  mutate(HerdDurableKey = as.factor(HerdDurableKey), 
         DairyYear = as.factor(DairyYear))

# which(duplicated(df_model$AnimalDurableCode)) # sanity check
cat("s34 dataset has", nrow(df_model), "events,",
    unique(df_model$AnimalDurableCode) %>% length(), "cows, and",
    unique(df_model$HerdDurableKey) %>% length(), "herds.")

glm_s34_on_year_milktype <- glm(s34~DairyYear+new_milk_type+HerdDurableKey+ 0, 
# add 2007 into the equation by +0, to make the intercept == OAD
                                binomial(link = "logit"), df_model)
s <- summary(glm_s34_on_year_milktype)
print(s$call)
# cat("deviance residuals:\n"); summary(s$deviance.resid)
coefs <- data.frame(s$coefficients[grepl("new_milk_type", rownames(s$coefficients)),])
coefs$sign <- ifelse(coefs$Pr...z..<0.001, "***", 
                     ifelse(coefs$Pr...z..<0.01, "**",
                            ifelse(coefs$Pr...z..<0.05, "*",
                                   ifelse(coefs$Pr...z..<0.1, ".",""))))
knitr::kable(coefs)
cat("\nAIC:",s$aic)
cat("\n##--ANOVA with Chi-sq test--##\n\n")
anovas <- anova(glm_s34_on_year_milktype, test = "Chisq")
print(anovas)

coefs$survival <- "s34"; anovas$survival <- "s34"
write.csv(coefs, paste0(outputDir, "coef_s34_milk_type.csv"), 
          row.names = T, quote = F)
write.csv(coefs, paste0(outputDir, "anova_s34_milk_type.csv"), 
          row.names = T, quote = F)

```


## S45

Transitioning herd at s12, s23 and s34 decreases the log odd ratio of the
survival. TAD decreases and transitioning at s45 increases the log odd ratio but
their effects are insignificant.

```{r s45_milk_type, cache = TRUE, eval = FALSE}
df_model <- select(df_downsample, AnimalDurableCode, HerdDurableKey, 
                   DairyYear, AgeParity,
                   new_milk_type, s45, starts_with("EBV_")) %>% 
  filter(AgeParity==5) %>% 
  distinct() %>% 
  # mutate(cg = paste0(HerdDurableKey, "_", DairyYear))
  mutate(HerdDurableKey = as.factor(HerdDurableKey), 
         DairyYear = as.factor(DairyYear))

# which(duplicated(df_model$AnimalDurableCode)) # sanity check
cat("s45 dataset has", nrow(df_model), "events,",
    unique(df_model$AnimalDurableCode) %>% length(), "cows, and",
    unique(df_model$HerdDurableKey) %>% length(), "herds.")

glm_s45_on_year_milktype <- glm(s45~DairyYear+new_milk_type+HerdDurableKey + 0, 
# add 2007 into the equation by +0, to make the intercept == OAD
                                binomial(link = "logit"), df_model)
s <- summary(glm_s45_on_year_milktype)
print(s$call)
# cat("deviance residuals:\n"); summary(s$deviance.resid)
coefs <- data.frame(s$coefficients[grepl("new_milk_type", rownames(s$coefficients)),])
coefs$sign <- ifelse(coefs$Pr...z..<0.001, "***", 
                     ifelse(coefs$Pr...z..<0.01, "**",
                            ifelse(coefs$Pr...z..<0.05, "*",
                                   ifelse(coefs$Pr...z..<0.1, ".",""))))
knitr::kable(coefs)
cat("\nAIC:",s$aic)
cat("\n##--ANOVA with Chi-sq test--##\n\n")
anovas <- anova(glm_s45_on_year_milktype, test = "Chisq")
print(anovas)

coefs$survival <- "s45"; anovas$survival <- "s45"
write.csv(coefs, paste0(outputDir, "coef_s45_milk_type.csv"), 
          row.names = T, quote = F)
write.csv(coefs, paste0(outputDir, "anova_s45_milk_type.csv"), 
          row.names = T, quote = F)

```

## Look at TOP (G)EBV within each milk type and their effects on 3 types of survivals

### choose EBV or GBV

```{r choose ebv or gbv}
bv <- "GBV" # EBV GBV
if(!bv %in% dir(outputDir)) dir.create(paste0(outputDir, bv, "/"), showWarnings = F)
```

### survival ~ cg + trait BV (within milk type)

```{r simple survival model, cache = TRUE}
rm(list=ls(pattern="idx_"))
# FUNCTION
stack_within_survival_type <- function(df_model, survival = "s12", 
                                  milk_type = "OAD") {
  
  t <- Sys.time()
  # out <- lapply(unique(df_model$new_milk_type), function(milk_type) {
  
  by_milk_type <- lapply(
    grep(bv, names(df_model), value = T), function(trait) {
cat(milk_type, trait, " ");print(Sys.time()-t)
      df <- filter(df_model, new_milk_type == milk_type)
      
      # all survived or all dead doesn't need modeling
      if(length(unique(df[[survival]])) < 2) {
        cat("All survived or dead\n")
        return(NULL)
      }
      
      # all EBVs are -9999
      if(length(unique(df[[trait]]))<=1) {
        cat("Missing EBV\n")
        return(NULL)
      }

     # fml <- as.formula(paste0(survival, "~DairyYear+HerdDurableKey+",trait))
      fml <- as.formula(paste0(survival, "~cg+",trait)) #, "+", trait, "*new_milk_type"))
      glm_s12_model <- glm(fml, 
                           gaussian(link="identity"), # binomial(link = "logit"), 
                           df)

      s <- summary(glm_s12_model)
      coefs <- data.frame(s$coefficients[grepl(bv, rownames(s$coefficients)),
                                         ,drop = F])
     names(coefs)[4] <- "p-value"
     coefs$sign <- ifelse(coefs$`p-value`<0.001, "***", 
                     ifelse(coefs$`p-value`<0.01, "**",
                            ifelse(coefs$`p-value`<0.05, "*",
                                   ifelse(coefs$`p-value`<0.1, ".",""))))

      coefs$milk_type <-  milk_type
      coefs$survival <- survival
      
      anovas <- anova(glm_s12_model, test = "Chisq")
      names(anovas)[5] <- "p-value"
      anovas$sign <- ifelse(anovas$`p-value`<0.001, "***", 
                           ifelse(anovas$`p-value`<0.01, "**",
                                  ifelse(anovas$`p-value`<0.05, "*",
                                         ifelse(anovas$`p-value`<0.1, ".",""))))
      anovas$milk_type <- milk_type
      anovas$survival <- survival
      
      return(list(coefs=coefs, anova = as.data.frame(anovas)))
      })
  
  coefs <- do.call(rbind, lapply(by_milk_type, `[[`, "coefs"))
  anovas <- do.call(rbind, lapply(by_milk_type, `[[`, "anova"))
cat("Finish stack_with_survival. Used "); print(Sys.time()-t)  
  return(list(coef = coefs, anova = anovas))
  # })
} # end function

# lapply will use up memories and crash
for(i in 2:5) {
cat("\n", paste0("s", i-1, i), "\n")
  df_model <- select(df_downsample, AnimalDurableCode, HerdDurableKey, 
                   DairyYear, AgeParity,cg, new_milk_type, 
                   matches(paste0("s", i-1, i)), 
                   starts_with(bv)) %>% 
  filter(AgeParity==i & 
           !grepl(paste0("Transition_s[", i,"-6]"), new_milk_type)) %>% 
  distinct() %>% 
  mutate(HerdDurableKey = as.factor(HerdDurableKey), 
         DairyYear = as.factor(DairyYear),
         cg = as.factor(cg),
         new_milk_type = as.factor(new_milk_type))
  
  for (j in unique(df_model$new_milk_type)) {
    assign(paste0("glm_s", i-1, i, "_", j),
           stack_within_survival_type(df_model, paste0("s", i-1, i), j))
    
    l <- get(paste0("glm_s", i-1, i, "_", j))
    if(length(l$coef) > 0) {
      write.csv(l$coef, paste0(outputDir, bv, "/coef_glm_s", i-1, i, "_", j, ".csv"),
              quote = F)
    }
    if(length(l$anova) >0) {
    write.csv(l$anova, paste0(outputDir, bv, "/anova_glm_s", i-1, i, "_", j, ".csv"),
              quote = F)
    }
    # saveRDS(get(paste0("glm_s", i-1, i, "_", j)), 
            # paste0(outputDir, "glm_s", i-1, i, "_", j, ".RData"))
  }
} # end do loop

# write.csv(coef, paste0(outputDir, "coef_s12_ebv.csv"), row.names = T, quote = F)
# write.csv(anovas, paste0(outputDir, "anova_s12_ebv.csv"), 
          # row.names = T, quote = F)
```

## Look at if EBV has an effect in different milking type

survival ~ cg + milk_type + trait EBV + trait EBV * milk_type
                                 
```{r interaction model}
rm(list=ls(pattern="idx_"))
# FUNCTION
stack_within_survival <- function(df_model, survival = "s12", 
                                  milk_type = "new_milk_type") {
  
  t <- Sys.time()
  # out <- lapply(unique(df_model$new_milk_type), function(milk_type) {
  
  by_milk_type <- lapply(
    grep(bv, names(df_model), value = T), function(trait) {
cat(trait, " ");print(Sys.time()-t)
      
      # all survived or all dead doesn't need modeling
      if(length(unique(df_model[[survival]])) < 2) {
        cat("All survived or dead\n")
        return(NULL)
      }

      # all EBVs are -9999
      if(length(unique(df_model[[trait]]))<=1) {
        cat("Missing EBV\n")
        return(NULL)
      }
      
      if(distribution=="gaussian") {
        fml <- as.formula(paste0(survival, "~cg+", milk_type, "+",trait,
                                 "+", trait, "*", milk_type, "-1"))
        glm_s12_model <- glm(fml, gaussian(link="identity"), df_model)
        
      } else {
        fml <- as.formula(paste0(survival, "~DairyYear+HerdDurableKey+",
                                      trait, "+", trait, "*", milk_type, "+0"))
        glm_s12_model <- glm(fml, binomial(link = "logit"), df_model)
      }

      s <- summary(glm_s12_model)
      coefs <- data.frame(s$coefficients[grepl(paste0(bv, "|new"), 
                                               rownames(s$coefficients)),
                                         ,drop = F])
     names(coefs)[4] <- "p-value"
     coefs$sign <- ifelse(coefs$`p-value`<0.001, "***", 
                     ifelse(coefs$`p-value`<0.01, "**",
                            ifelse(coefs$`p-value`<0.05, "*",
                                   ifelse(coefs$`p-value`<0.1, ".",""))))

      coefs$survival <- survival
      
      anovas <- anova(glm_s12_model, test = "Chisq")
      names(anovas)[5] <- "p-value"
      anovas$sign <- ifelse(anovas$`p-value`<0.001, "***", 
                           ifelse(anovas$`p-value`<0.01, "**",
                                  ifelse(anovas$`p-value`<0.05, "*",
                                         ifelse(anovas$`p-value`<0.1, ".",""))))
      anovas$survival <- survival
      
      return(list(coefs=coefs, anova = as.data.frame(anovas)))
      })
  
  coefs <- do.call(rbind, lapply(by_milk_type, `[[`, "coefs"))
  anovas <- do.call(rbind, lapply(by_milk_type, `[[`, "anova"))
cat("Finish stack_with_survival. Used "); print(Sys.time()-t)  
  return(list(coef = coefs, anova = anovas))
  # })
} # end function

# lapply will use up memories and crash
for(i in 2:5) {
cat("\n", paste0("s", i-1, i), "\n")
  df_model <- select(df_downsample, AnimalDurableCode, HerdDurableKey, 
                   DairyYear, AgeParity,cg, new_milk_type, 
                   matches(paste0("s", i-1, i)), 
                   starts_with(bv)) %>% 
  filter(AgeParity==i & 
           !grepl(paste0("Transition"), new_milk_type)) %>% # only OAD and TAD
  distinct() %>% 
  mutate(HerdDurableKey = as.factor(HerdDurableKey), 
         DairyYear = as.factor(DairyYear),
         cg = as.factor(cg),
         new_milk_type = as.factor(new_milk_type))
  
  # Make TAD as the reference type
  levels(df_model$new_milk_type) <- c("TAD", "OAD")
  
    assign(paste0("glm_s", i-1, i),
           stack_within_survival(df_model, paste0("s", i-1, i)))
    
    l <- get(paste0("glm_s", i-1, i))
    if(length(l$coef) > 0) {
      write.csv(l$coef, paste0(outputDir, bv, "/coef_glm_x_s", i-1, i, ".csv"),
              quote = F)
    }
    if(length(l$anova) >0) {
    write.csv(l$anova, paste0(outputDir, bv, "/anova_glm_x_s", i-1, i, ".csv"),
              quote = F)
    }
} # end do loop

```

```{r, eval = F}
rmarkdown::render(output_dir = outputDir, 
                  input = rstudioapi::getSourceEditorContext()$path)
```

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold', eval=FALSE}

```
