---
title: "Post modeling analysis"
author: "Luna Zhang"
date: "`r Sys.Date()`"
output: #> rmarkdown::html_vignette
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F, warning = F, message = F,
                      root.dir = "e:/500390_OAD/output/modeling/"
)
```

Modelling of survival on effects.

```{r, include =FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)
# library(plm)

source("utils.R")


outputDir <- "C:/Users/lzhang/AbacusBio Ltd/Pivot Cloud Drive - 500390 - DNZ TOPs impact on OAD survival 2021-3637/2. Doing/Results/" # "c:/Users/lzhang/Documents/GitHub/500390_OAD/output/" # "z:/Luna/500390_OAD/explore/"
# rawfileDir <- "e:/OAD Dairy Data/2022021110_EBVs/"
fileDir <-  "e:/500390_OAD/output/modeling/"

# if(!exists(outputDir)) dir.create(outputDir)
```

```{r load, cache=TRUE, cache.lazy=FALSE, cache.extra=tools::md5sum("e:/500390_OAD/output/modeling/df_downsample.RData")}
df_downsample <- readRDS(paste0(fileDir, "df_downsample.RData"))

```

# Sampled data summary statistics

## Herd size and cow count across years

```{r}
cat("Downsampled dataset has", nrow(df_downsample), "events,",
    unique(df_downsample$AnimalDurableCode) %>% length(), "cows, and",
    unique(df_downsample$HerdDurableKey) %>% length(), "herds.")

group_by(df_downsample, herd_milk_type, HerdDurableKey) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% group_by(herd_milk_type) %>% 
  tally() %>% 
  mutate(prop = n/length(unique(df_downsample$HerdDurableKey))) %>% 
  knitr::kable(align = "l", 
               caption = "Total N herds by milking type across years",
               digits = 2) %>% 
  kable_styling(full_width = T)

select(df_downsample, new_milk_type, AnimalDurableCode) %>% 
  distinct() %>% 
  group_by(new_milk_type) %>% 
  tally() %>% ungroup() %>% 
  mutate(prop = n/length(unique(df_downsample$AnimalDurableCode))) %>% 
  knitr::kable(align = "l",
               caption = "Total N animals by milking type across years", 
               digits = 3,
               format.args = list(big.mark=",")) %>% 
  kable_styling(full_width = T)
```

## count by age parity

Herd size by age parity

```{r herd_size_by_age}

group_by(df_downsample, herd_milk_type, AgeParity, HerdDurableKey) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% group_by(herd_milk_type, AgeParity) %>% 
  tally() %>% ungroup() %>% 
  mutate(prop = n/sum(n), AgeParity = as.factor(AgeParity)) %>% 
  # knitr::kable(align = "l", 
  #              caption = "N herds by herd Regime type and Age parity", digits = 2) %>% 
  # kable_styling(full_width = T)
  ggplot(aes(herd_milk_type, n, fill = AgeParity)) +
  geom_bar(stat = "identity", width = .8, color = "black", 
           position = position_dodge(.8)) +
  labs(x = "") +
  theme_pubclean() +
  scale_fill_brewer(palette = "Set3")
```

Number of cows by age parity

```{r n_cows_by_age}
select(df_downsample, AgeParity, new_milk_type, AnimalDurableCode) %>% 
  distinct() %>% 
  group_by(new_milk_type, AgeParity) %>% 
  tally() %>% ungroup() %>% 
  # group_by(AgeParity, new_milk_type, AnimalDurableCode) %>% 
  mutate(prop = n/sum(n), AgeParity = as.factor(AgeParity)) %>% 
  # knitr::kable(align = "l",
               # caption = "N animals by milking type", digits = 3,
               # format.args = list(big.mark=",")) %>% 
  # kable_styling(full_width = T)
  ggplot(aes(new_milk_type, n, fill = AgeParity)) +
  geom_bar(stat = "identity", width = .8, color = "black", 
           position = position_dodge(.8)) +
  scale_y_continuous(breaks = c(0, 5000, 10000, 50000, 100000, 150000)) +
  labs(x = "") +
  theme_pubclean() +
  scale_fill_brewer(palette = "Set3")

```
## herd size and n cow for paper

```{r count_for_paper}
# herd size

df_sub <- filter(df_downsample, herd_milk_type %in% c("TAD", "OAD")) 
  
herd_age_type <- 
  group_by(df_sub, herd_milk_type, AgeParity, HerdDurableKey) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% group_by(herd_milk_type, AgeParity) %>% 
  tally() %>% ungroup() %>% 
  mutate(prop = n/length(unique(df_sub$HerdDurableKey))*100,
         AgeParity = as.factor(AgeParity))
names(herd_age_type)[3:4] <- c("n_herd", "prop_herd")

# herd_age_type %>% knitr::kable(digits = 2) 

cow_age_type <-
  select(df_sub, AgeParity, herd_milk_type, AnimalDurableCode) %>% 
  distinct() %>% 
  group_by(herd_milk_type, AgeParity) %>% 
  tally() %>% ungroup() %>% 
  mutate(prop = n/length(unique(df_sub$AnimalDurableCode))*100, 
         AgeParity = as.factor(AgeParity))
names(cow_age_type)[3:4] <- c("n_cow", "prop_cow")

# cow_age_type %>% knitr::kable(digits = 2)

age_type <- full_join(herd_age_type, cow_age_type)
write.csv(age_type, paste0(outputDir, "n_herd_cow_by_age.csv"), quote = F,
          row.names = F)

age_type %>% knitr::kable(digits = 2) #%>% kable_styling()
```


```{r essential_traits}
essential_traits <- c("BodyConditionScore", "Fertility_CR42", 
                      "FunctionalSurvival", "SomaticCellScore",
                      "TotalVolume", "TotalFat", "TotalProtein",
                      "TOP_UdderSupport", "TOP_MilkingSpeed", "TOP_Capacity", 
                      "TOP_FrontTeat" )
```

# Simple model within each survival and milk_type levels

e.g. `glm(formula = s12 ~ cg + Trait, family = gaussian(link = "identity"),  ...)`

```{r read_simple_model}

test <- lapply(c("EBV", "GBV"), function(bv) {
  coef_glms <- lapply(dir(paste0(fileDir, bv, "/"), "coef_glm_s"), function(i) {
    return(read.csv(paste0(fileDir, bv, "/", i), header = T) %>% 
           rename(Trait = X) %>% 
          mutate(sign = ifelse(is.na(sign), "", 
                                ifelse(sign==".", ".", sign)),
                  predictor = Trait,
                 evaluation = bv))
    })

  anova_glms <- lapply(dir(paste0(fileDir, bv, "/"), "anova_glm_s"), function(i) {
    return(read.csv(paste0(fileDir, bv, "/", i), header = T) %>% 
           rename(Trait = X) %>% 
           mutate(sign = ifelse(is.na(sign), "", 
                                ifelse(sign==".", ".", sign)),
                  predictor = Trait,
                  evaluation = bv))
    })

  df_coef_glms <- do.call(rbind, coef_glms)
  df_anova_glms <- do.call(rbind, anova_glms)
  
  # change trait
  df_coef_glms$Trait <- gsub(paste0(bv, "_"), "", df_coef_glms$Trait)
  df_coef_glms$predictor <- gsub(paste0(bv, "_"), "", df_coef_glms$predictor)
  df_anova_glms$Trait <- gsub(paste0(bv, "_"), "", df_anova_glms$Trait)
  df_anova_glms$predictor <- gsub(paste0(bv, "_"), "", df_anova_glms$predictor)

  if(grepl("Chi", names(df_anova_glms)) %>% sum()>0) {
    names(df_anova_glms)[5] <- "Resid..Dev"
    names(df_anova_glms)[6] <- "p-value"
    df_anova_glms$sign <- ifelse(df_anova_glms$`p-value`<0.001, "***", 
                           ifelse(df_anova_glms$`p-value`<0.01, "**",
                                  ifelse(df_anova_glms$`p-value`<0.05, "*",
                                         ifelse(df_anova_glms$`p-value`<0.1, ".",""))))
  
    df_anova_glms <- df_anova_glms %>% 
      select(Trait, Df, Deviance, Resid..Df, Resid..Dev, `p-value`, sign, milk_type,
             survival)
  }
  return(list(coef =  df_coef_glms, anova = df_anova_glms))
})

names(test) <- c("ebv", "gbv")

df_coef_glms_EBV <- test$ebv$coef
df_coef_glms_GBV <- test$gbv$coef

df_anova_glms_EBV <- test$ebv$anova
df_anova_glms_GBV <- test$gbv$anova
  
# df_anova_glms <- df_anova_glms %>% 
#            mutate(sign = ifelse(is.na(sign), "", sign))

```

#> ## OAD only, Coefficients for all traits' EBVs

```{r, echo=FALSE, eval=FALSE}
# CONVERT COEF OUTPUT
print_coef <- select(df_coef_glms_EBV, -matches("value")) %>% 
  pivot_wider(names_from = milk_type, values_from = c(Estimate, Std..Error, sign),
               names_glue = "{milk_type}_{.value}") %>% 
  select(survival, Trait, matches("OAD"), matches("TAD"), 
         matches("s12"), matches("s23"), matches("s34"), matches("s45")) %>% 
  mutate(across(where(is.numeric), ~ifelse(.x==round(.x),
    round(.x),formatC(.x, digits = 2, format = "fg"))
    ))

select(print_coef, survival, Trait, matches("OAD")) %>% 
  knitr::kable(format.args = list(big.mark=","), align = "l",
             caption = "Linear regression coefficient statistics") %>% 
  kable_styling(fixed_thead = T)
```

## Coefficients for all trait EBVs within milking types

### EBV and GEBV

#### Essential traits

##### TAD and OAD only

```{r simple_model_coef, fig.show='asis', fig.height=16, fig.width=8}
invisible( # avoind printing lapply outputs e.g. [1][[1]] [2][[2]]
  lapply(c("EBV", "GBV"), function(bv) {
 # lapply(seq(essential_traits), function(i) {
    
    df <- get(paste0("df_coef_glms_", bv)) %>% 
          #   filter(Trait == essential_traits[i]) %>% 
        filter(Trait %in% essential_traits) %>% 
             mutate(milk_type = 
                      factor(milk_type, 
                             levels = c("TAD", setdiff(unique(milk_type), "TAD")
                                        ))) %>% 
      filter(as.integer(milk_type)<=2)
   # color <- ifelse(i==length(essential_traits), "black", "white")
    
    p <- ggplot(#get(paste0("df_coef_glms_", bv)) %>%
           #filter(Trait %in% essential_traits),
      df,
         aes(survival, Estimate, fill = milk_type)) +
    geom_bar(stat="identity", color="black", position = position_dodge(.9)) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position=position_dodge(.9)) +
      # ylim(min(c(min(df$Estimate)-max(abs(df$Std..Error)), 0)),
             # max(df$Estimate)+max(abs(df$Std..Error))) +
               #diff(range(df$Estimate+df$Std..Error))/100 +
    geom_text(aes(y = Estimate+Std..Error, label = sign),
              position = position_dodge(.9)
              #, vjust = -diff(range(df$Estimate))/100
              ) +
    labs(title = bv #ifelse(i==1, paste0(bv, " ", essential_traits[i]),
                        # essential_traits[i]),
         # x = "", y = ""
         ) +
    facet_grid(Trait~., scales = "free_y") +
    # scale_fill_brewer(palette = "Set3") +
    scale_fill_manual(values =
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(df$milk_type)-1, "Set3"))) +
    theme_pubclean() # +
      # theme(legend.position = ifelse(i==1, "top", "none"),
            # axis.text.x = element_text(color = color)
            # )
    print(p)
 # })
})
)

```

*Foot notes for the signs of P-values:*  
\*\*\*: $p<0.001$  
\*\*: $0.001 \le p<0.01$  
\*: $0.01 \le p<0.05$  
.: $0.05 \le p<0.1$  

##### All 5 milking types

```{r simple_model_coef1, fig.show='asis', fig.height=16, fig.width=8}

invisible(
  lapply(c("EBV", "GBV"), function(bv) {
 # lapply(seq(essential_traits), function(i) {
    
    df <- get(paste0("df_coef_glms_", bv)) %>% 
          #   filter(Trait == essential_traits[i]) %>% 
        filter(Trait %in% essential_traits) %>% 
             mutate(milk_type = 
                      factor(milk_type, 
                             levels = c("TAD", setdiff(unique(milk_type), "TAD")
                                        )))
   # color <- ifelse(i==length(essential_traits), "black", "white")
    
    p <- ggplot(#get(paste0("df_coef_glms_", bv)) %>%
           #filter(Trait %in% essential_traits),
      df,
         aes(survival, Estimate, fill = milk_type)) +
    geom_bar(stat="identity", color="black", position = position_dodge(.9)) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position=position_dodge(.9)) +
      # ylim(min(c(min(df$Estimate)-max(abs(df$Std..Error)), 0)),
             # max(df$Estimate)+max(abs(df$Std..Error))) +
               #diff(range(df$Estimate+df$Std..Error))/100 +
    geom_text(aes(y = Estimate+Std..Error, label = sign),
              position = position_dodge(.9)
              #, vjust = -diff(range(df$Estimate))/100
              ) +
    labs(title = bv #ifelse(i==1, paste0(bv, " ", essential_traits[i]),
                        # essential_traits[i]),
         # x = "", y = ""
         ) +
    facet_grid(Trait~., scales = "free_y") +
    # scale_fill_brewer(palette = "Set3") +
    scale_fill_manual(values =
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(df$milk_type)-1, "Set3"))) +
    theme_pubclean() # +
      # theme(legend.position = ifelse(i==1, "top", "none"),
            # axis.text.x = element_text(color = color)
            # )
    print(p)
 # })
})
)
```

#### Other traits

Transformation of nonlinear trait evaluations do not seems to have more power
in the model (new value = old value + old value^2)

```{r simple_model_coef_other, fig.show='asis', fig.height=60, fig.width=8}

invisible(
  lapply(c("EBV", "GBV"), function(bv) {
 # lapply(seq(essential_traits), function(i) {
    
    df <- get(paste0("df_coef_glms_", bv)) %>% 
          #   filter(Trait == essential_traits[i]) %>% 
        filter(!Trait %in% essential_traits) %>% 
             mutate(milk_type = 
                      factor(milk_type, 
                             levels = c("TAD", setdiff(unique(milk_type), "TAD")
                                        )))
   # color <- ifelse(i==length(essential_traits), "black", "white")
    
    p <- ggplot(#get(paste0("df_coef_glms_", bv)) %>%
           #filter(Trait %in% essential_traits),
      df,
         aes(survival, Estimate, fill = milk_type)) +
    geom_bar(stat="identity", color="black", position = position_dodge(.9)) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position=position_dodge(.9)) +
      # ylim(min(c(min(df$Estimate)-max(abs(df$Std..Error)), 0)),
             # max(df$Estimate)+max(abs(df$Std..Error))) +
               #diff(range(df$Estimate+df$Std..Error))/100 +
    geom_text(aes(y = Estimate+Std..Error, label = sign),
              position = position_dodge(.9) #,
              #, vjust = -diff(range(df$Estimate))/100
              ) +
    labs(title = bv #ifelse(i==1, paste0(bv, " ", essential_traits[i]),
                        # essential_traits[i]),
         # x = "", y = ""
         ) +
    facet_grid(Trait~., scales = "free_y") +
    # scale_fill_brewer(palette = "Set3") +
    scale_fill_manual(values =
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(df$milk_type)-1, "Set3"))) +
    theme_pubclean() # +
      # theme(legend.position = ifelse(i==1, "top", "none"),
            # axis.text.x = element_text(color = color)
            # )
    print(p)
 # })
})
)
```

### Comparison EBV vs GEBV

#### Essential traits 

```{r coef_ebv_gbv_simple, fig.width=8, fig.height=16}

df <- rbind(df_coef_glms_EBV, df_coef_glms_GBV)
write.csv(df, paste0(outputDir, "coef_glm_all.csv"))

df_sub <- filter(df, Trait %in% essential_traits) %>% 
  mutate(milk_type = 
           factor(milk_type, 
                  levels = c("TAD", setdiff(unique(milk_type), "TAD"))))

ggplot(df_sub, aes(x = survival, y = Estimate, fill = evaluation)) + 
    geom_bar(stat = "identity", color = "black", 
             position = position_dodge()) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position = position_dodge(.9)) +
    geom_text(aes(y = Estimate+Std..Error, label = sign), 
              position = position_dodge(.9)) +
    facet_grid(Trait~milk_type, scales = "free") +
  scale_fill_brewer(palette = "Paired") +
  theme_pubclean() # +
```

#### Other traits 

```{r coef_ebv_gbv_simple_other, fig.width=8, fig.height=60}

df_sub <- filter(df, !Trait %in% essential_traits) %>% 
  mutate(milk_type = 
           factor(milk_type, 
                  levels = c("TAD", setdiff(unique(milk_type), "TAD"))))

ggplot(df_sub, aes(x = survival, y = Estimate, fill = evaluation)) + 
    geom_bar(stat = "identity", color = "black", 
             position = position_dodge()) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position = position_dodge(.9)) +
    geom_text(aes(y = Estimate+Std..Error, label = sign), 
              position = position_dodge(.9)) +
    facet_grid(Trait~milk_type, scales = "free") +
  scale_fill_brewer(palette = "Paired") +
  theme_pubclean() # +
```

## Model powers

```{r, eval=FALSE}

# convert anova output
print_anova <- filter(df_anova_glms, grepl("EBV", Trait)) %>% 
  mutate(R2 = Deviance/(Resid..Dev+Deviance)) %>% 
  select(Trait, R2, `p-value`, sign, milk_type, survival) %>% 
  pivot_wider(id_cols = matches("survival|Trait"), 
              names_from = milk_type, values_from = c(R2, sign), 
              names_glue = "{milk_type}_{.value}") %>% 
  select(survival, Trait, matches("OAD"), matches("TAD"), 
         matches("s12"), matches("s23"), matches("s34"), matches("s45")) %>% 
  mutate(across(where(is.numeric), ~ifelse(.x==round(.x),
    round(.x),formatC(.x, digits = 2, format = "fg"))))

knitr::kable(print_anova, digits = 3, align = "l",
             caption = "Linear regression model statistics (ANOVA+Chi-sq)" )%>% 
  kable_styling(fixed_thead = T)

# write.csv(print_coef, paste0(fileDir, "coef_glm_all.csv"), row.names = F, quote = F)
# write.csv(print_anova, paste0(fileDir, "anova_glm_all.csv"), row.names = F, quote = F)
```

# Trait and milk_type interaction model

e.g. `glm(formula = s12 ~ cg + Trait + milk_type + Trait*milk_type, family = gaussian(link = "identity"),  ...)`  
Only look at OAD and TAD. Each model takes up to 10 min to run.  
The objective is to see if the trait EBV has an effect and if its effects are
different in different milking types (OAD and TAD).

```{r read_interaction_model}
test <- lapply(c("EBV", "GBV"), function(bv) {
  coef_glms <- lapply(dir(paste0(fileDir, bv, "/"), "coef_glm_x_s"), function(i) {
    return(read.csv(paste0(fileDir, bv, "/", i), header = T) %>%
           rename(Trait = X) %>%
           mutate(sign = ifelse(is.na(sign), "", 
                                ifelse(sign==".", ".", sign)),
                  predictor = Trait))
    })
  
  anova_glms <- lapply(dir(paste0(fileDir, bv, "/"), "anova_glm_x_s"), 
                       function(i) {
    return(read.csv(paste0(fileDir, bv, "/", i), header = T) %>% 
             rename(Trait = X) %>% 
           mutate(sign = ifelse(is.na(sign), "", 
                                ifelse(sign==".", ".", sign)),
                  predictor = Trait))
    })

  df_coef_glms <- do.call(rbind, coef_glms)
  df_anova_glms <- do.call(rbind, anova_glms)
  
  # change Trait
  idx <- grep(paste0("^", bv), df_anova_glms$predictor)
  idx1 <- grep("NULL", df_anova_glms$predictor)
  d <- diff(idx1) %>% unique()
  for(i in seq(idx1)) {
    df_anova_glms$Trait[idx1[i]:(idx1[i]+d-1)] <- df_anova_glms$predictor[idx[i]]
  }
  df_anova_glms$Trait <- gsub(paste0(bv, "_"), "", df_anova_glms$Trait)

  idx <- grep("^new", df_coef_glms$predictor)
  d <- diff(idx) %>% unique()
  for(i in seq(idx)) {
    df_coef_glms$Trait[(idx[i]-d+1):idx[i]] <- df_coef_glms$predictor[idx[i]-d+1]
  }
  df_coef_glms$Trait <- gsub(paste0(bv, "_"), "", df_coef_glms$Trait)
  
  # change predictor
  df_anova_glms$predictor[df_anova_glms$predictor=="new_milk_type"] <- 
    "milking_type"
  df_anova_glms$predictor[df_anova_glms$predictor==
                            paste0(bv, "_", df_anova_glms$Trait)] <- "Trait"
  df_anova_glms$predictor[grepl(":", df_anova_glms$predictor)] <- 
    "Trait_x_milking_type"
  
  idx <- seq(1, nrow(df_coef_glms), by = 2) # odd value
  df_coef_glms$predictor[idx] <- "Trait"
  df_coef_glms$predictor[idx+1] <- "Trait_x_OAD"
  
  return(list(coef = df_coef_glms, anova = df_anova_glms))
})
names(test) <- c("ebv", "gbv")

df_coef_glms_x_EBV <- test$ebv$coef
df_coef_glms_x_GBV <- test$gbv$coef

df_anova_glms_x_EBV <- test$ebv$anova
df_anova_glms_x_GBV <- test$gbv$anova
```

## Coefficients

In the model, the contrast levels are cg, TAD, OAD, trait\*TAD and trait\*OAD.
The reference (intercept) is TAD cows, i.e., TAD=0 and OAD=1, so
TAD+trait\*TAD=0. OAD effect isn't in the modeling function output, probably
because OAD effect by itself is hard to interpret.
Therefore, only the coefficients of Trait and Trait\*OAD are used to test 
against the reference (CG effect omit here). 

For example, if s12=$\mu$+0.03\*functional survival+0.01\*functional survival
\*milking type (CG effect omit here), then it means when the milking type is TAD,
the effect of functional survival is 0.03, and s12=$\mu$+0.03. However, when the 
milking type is OAD, the effect of functional survival is 0.03+0.01, and s12=
$\mu$+0.04.

The stars mean if their effects are significant in the model.

### EBV and GBV

#### Essential traits

```{r coef, fig.width=8, fig.height=16}
# bv <- "EBV"

invisible(lapply(c("EBV", "GBV"), function(bv) {
  p <- ggplot(get(paste0("df_coef_glms_x_", bv)) %>% 
                filter(Trait %in% essential_traits), 
              aes(x = survival, y = Estimate, fill = predictor)) + 
    geom_bar(stat = "identity", color = "black", 
             position = position_dodge()) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position = position_dodge(.9)) +
    geom_text(aes(y = Estimate+Std..Error, label = sign), 
              position = position_dodge(.9)) +
    labs(title = bv) +
    facet_grid(Trait~., scales = "free_y") +
  scale_fill_brewer(palette = "Set3") +
  theme_pubclean() # +
    # theme(legend.position = "none")
print(p)
}))


```

#### Other traits

Transformation of nonlinear trait evaluations do not seems to have more power
in the model (new value = old value + old value^2)

```{r coef_other, fig.width=8, fig.height=60}
# bv <- "EBV"

invisible(lapply(c("EBV", "GBV"), function(bv) {
  p <- ggplot(get(paste0("df_coef_glms_x_", bv)) %>% 
                filter(!Trait %in% essential_traits), 
              aes(x = survival, y = Estimate, fill = predictor)) + 
    geom_bar(stat = "identity", color = "black", 
             position = position_dodge()) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position = position_dodge(.9)) +
    geom_text(aes(y = Estimate+Std..Error, label = sign), 
              position = position_dodge(.9)) +
    labs(title = bv) +
    facet_grid(Trait~., scales = "free_y") +
  scale_fill_brewer(palette = "Set3") +
  theme_pubclean() # +
    # theme(legend.position = "none")
print(p)
}))
```


```{r coef_table, echo=FALSE, eval=FALSE}
 
# convert coef output
print_coef <- df_coef_x_glms %>% 
  select(Trait, predictor, Estimate, sign, survival) %>% 
  pivot_wider(id_cols = matches("Trait|predi"), 
              names_from = survival, values_from = c(Estimate, sign), 
              names_glue = "{survival}_{.value}") %>% 
  select(Trait, predictor, matches("s12"), matches("s23"), 
         matches("s34"), matches("s45")) %>% 
  mutate(across(where(is.numeric), ~ifelse(.x==round(.x),
    round(.x), formatC(.x, digits = 2, format = "fg"))))

knitr::kable(print_coef, align = "l",
             format.args = list(big.mark = ","), 
             caption = "Linear regression model coeficient statistics" )%>% 
  kable_styling(fixed_thead = T)

# write.csv(print_coef, paste0(fileDir, "coef_glm_x_all.csv"), 
          # row.names = F, quote = F)
# write.csv(print_coef, paste0(fileDir, "coef_glm_x_all.csv"), 
          # row.names = F, quote = F)
```

### Comparison EBV vs GEBV

#### Essential traits

```{r coef_ebv_gbv, fig.width=8, fig.height=16}

df_coef_glms_x_EBV$evaluation <- "EBV"
df_coef_glms_x_GBV$evaluation <- "GEBV"

df <- rbind(df_coef_glms_x_EBV, df_coef_glms_x_GBV) # %>% 
 # filter(predictor == "Trait_x_OAD")
write.csv(df, paste0(outputDir, "coef_glm_x_all.csv"))

ggplot(filter(df, Trait %in% essential_traits),
       aes(x = survival, y = Estimate, fill = evaluation)) + 
    geom_bar(stat = "identity", color = "black", 
             position = position_dodge()) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position = position_dodge(.9)) +
    geom_text(aes(y = Estimate+Std..Error, label = sign), 
              position = position_dodge(.9)) +
    facet_grid(Trait~predictor, scales = "free_y") +
  scale_fill_brewer(palette = "Paired") +
  theme_pubclean() # +
```

#### Other traits

```{r coef_ebv_gbv_other, fig.width=8, fig.height=60}

ggplot(filter(df, !Trait %in% essential_traits),
       aes(x = survival, y = Estimate, fill = evaluation)) + 
    geom_bar(stat = "identity", color = "black", 
             position = position_dodge()) +
    geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  width=.2, position = position_dodge(.9)) +
    geom_text(aes(y = Estimate+Std..Error, label = sign), 
              position = position_dodge(.9)) +
    facet_grid(Trait~predictor, scales = "free_y") +
  scale_fill_brewer(palette = "Paired") +
  theme_pubclean() # +
```

#### For paper

#### Essential traits

##### TAD and OAD only

```{r simple_model_coef, fig.show='asis', fig.height=16, fig.width=10}
# cOMBINE THE P VALUE OF interaction model with single trait model picture
df_coef_glms_EBV <- mutate(df_coef_glms_EBV,
                           std_coef = Estimate*`Std..Error`)

test <- df_coef_glms_x_EBV %>% 
  filter(predictor == "Trait_x_OAD") %>% 
  mutate(sign_x = sign) %>% 
  select(Trait, survival, sign_x)

test <- left_join(df_coef_glms_EBV, test)
test$sign_x[test$milk_type=="TAD"] <- ""

# invisible( # avoind printing lapply outputs e.g. [1][[1]] [2][[2]]
#   lapply(c("EBV"), function(bv) {
 # lapply(seq(essential_traits), function(i) {
    bv <- "EBV"
    df <- test %>% 
          #   filter(Trait == essential_traits[i]) %>% 
        filter(Trait %in% essential_traits) %>% 
             mutate(milk_type = 
                      factor(milk_type, 
                             levels = c("TAD", setdiff(unique(milk_type), "TAD")
                                        ))) %>% 
      filter(as.integer(milk_type)<=2)
   # color <- ifelse(i==length(essential_traits), "black", "white")
    
    p <- ggplot(#get(paste0("df_coef_glms_", bv)) %>%
           #filter(Trait %in% essential_traits),
      df,
         aes(survival, std_coef, fill = milk_type)) +
    geom_bar(stat="identity", color="black", position = position_dodge(.9)) +
    # geom_errorbar(aes(ymin = Estimate-Std..Error, ymax = Estimate+Std..Error),
                  # width=.2, position=position_dodge(.9)) +
    geom_text(aes(y = std_coef, label = sign),
              position = position_dodge(.9)
              #, vjust = -diff(range(df$Estimate))/100
              ) +
    geom_text(aes(y = std_coef, label = sign_x),
              position = position_dodge(.9), color = "red"
              , vjust = 1
              ) +
    # labs(title = bv #ifelse(i==1, paste0(bv, " ", essential_traits[i]),
                        # essential_traits[i]),
         # x = "", y = ""
         # ) +
    facet_grid(Trait~., scales = "free_y") +
    # scale_fill_brewer(palette = "Set3") +
    scale_fill_manual(values =
                      c("grey", #ggpubr::get_palette("npg", 5)
                        RColorBrewer::brewer.pal(
                          nlevels(df$milk_type)-1, "Set3"))) +
    theme_pubclean() # +
      # theme(legend.position = ifelse(i==1, "top", "none"),
            # axis.text.x = element_text(color = color)
            # )
    print(p)
 # })
# })
# )

```


## Model powers

This is an ANOVA and Chi-sq test of the model predictors, to show if the trait
and interaction terms are significantly different.

Because new_milk_type is always insignificant in all models, the result isn't
shown in the table.

```{r, eval=FALSE,echo=FALSE}
# convert anova output
print_anova <- filter(df_anova_glms, !grepl("NULL|cg|type[0-9]*$", predictor)) %>% 
  mutate(R2 = Deviance/(Resid..Dev+Deviance)) %>% 
  select(Trait, predictor, R2, sign, survival) %>% 
  pivot_wider(id_cols = matches("Trait|predi"), 
              names_from = survival, values_from = c(R2, sign), 
              names_glue = "{survival}_{.value}") %>% 
  select(Trait, predictor, matches("s12"), matches("s23"), 
         matches("s34"), matches("s45")) %>% 
  mutate(across(where(is.numeric), ~ifelse(.x==round(.x),
    round(.x), formatC(.x, digits = 2, format = "fg"))))

knitr::kable(print_anova, align = "l",
             format.args = list(big.mark = ","), 
             caption = "Linear regression model statistics (ANOVA+Chi-sq)" )%>% 
  kable_styling(fixed_thead = T)

# write.csv(print_coef, paste0(fileDir, "coef_glm_x_all.csv"), 
          # row.names = F, quote = F)
# write.csv(print_anova, paste0(fileDir, "anova_glm_x_all.csv"), 
          # row.names = F, quote = F)
```

```{r, echo=FALSE, eval=FALSE}
knitr::kable(print_coef, align = "l",
             format.args = list(big.mark=","),
             caption = "Linear regression coefficient statistics") %>% 
  kable_styling(fixed_thead = T)
```


```{r, eval = F}
rmarkdown::render(output_dir = outputDir, 
                  input = rstudioapi::getSourceEditorContext()$path)
```